Load libraries
```{r}
library(biobroom)
library(cowplot)
library(RColorBrewer)
library(tidyverse)
source("../../General/rmb_functions.R")
source("../../General/parameters.R")
```

Load data
```{r}
#Mapping files
mb.map <- readRDS("../Data/GNOTO_MB_map.RDS") %>% filter(Compartment == "ES", Treatment == "Ino")
rn.map <- readRDS("../Data/RNASeq_map.RDS") %>% filter(Tissue == "Root" & Treatment == "Ino")

#Z-transformed matrices
mb.mtx <- readRDS("../Data/zscore_otu_mtx.RDS")
rn.mtx <- readRDS("../Data/zscore_gene_mtx.RDS")

#Clustering results
mb.clust <- readRDS("../Data/otu_clusters.RDS")
rn.clust <- readRDS("../Data/gene_clusters.RDS")

#Taxonomy
tax <- readRDS("../Data/SILVAcr_tax.RDS")

#OTU table
otu <- readRDS("../Data/SILVAcr_otu.RDS")
otu <- otu[rowSums(otu) > 0.05*ncol(otu),]
otu <- otu[,colnames(otu) %in% colnames(mb.mtx)]

#Core microbiome IDs
core.ids <- readRDS("../Data/core_otu_ids.RDS")
```

Perform single value decomposition on microbiome clusters
```{r}
#Generate lists to hold results
mb.svd.var.list <- list()
mb.svd.vec.list <- list()

#Perform SVD
for(cluster.id in mb.clust$Cluster %>% unique()){
  cluster.members <- filter(mb.clust, Cluster == cluster.id)$OTU_ID
  cluster.mtx <- mb.mtx[match(cluster.members, row.names(mb.mtx)),]
  svd(cluster.mtx)
  mb.svd.var.list[[cluster.id]] <- svd(cluster.mtx)$d^2/sum(svd(cluster.mtx)$d^2)
  mb.svd.vec.list[[cluster.id]] <-svd(cluster.mtx)$v
}

#Recover variance results and reformat
mb.svd.var.df <- plyr::ldply(mb.svd.var.list, function(x) x)
names(mb.svd.var.df) <- c("Cluster", paste("SVD", 1:(ncol(mb.svd.var.df) - 1), sep = ""))
mb.svd.var.df <- mb.svd.var.df %>% 
  gather(key = "SVD", value = "Variance", -Cluster) %>% 
  mutate(Variance = Variance * 100) 

#Recover singular vectors and reformat
mb.svd.vec.df <- plyr::ldply(mb.svd.vec.list, function(x) x)
names(mb.svd.vec.df) <- c("Cluster", paste("SVD", 1:(ncol(mb.svd.vec.df) - 1), sep = ""))
mb.svd.vec.df$SampleID <- colnames(mb.mtx)
mb.svd.vec.df <- mb.svd.vec.df %>% 
  gather(key = "SVD", value = "Score", -SampleID, -Cluster) %>% 
  inner_join(mb.map, by = "SampleID") %>% 
  inner_join(mb.svd.var.df, by = c("Cluster", "SVD"))
```

Perform single value decomposition on RNASeq clusters
```{r}
#Generate lists to hold results
rn.svd.var.list <- list()
rn.svd.vec.list <- list()

#Perform SVD
for(cluster.id in rn.clust$Cluster %>% unique()){
  cluster.members <- filter(rn.clust, Cluster == cluster.id)$gene
  cluster.mtx <- rn.mtx[match(cluster.members, row.names(rn.mtx)),]
  svd(cluster.mtx)
  rn.svd.var.list[[cluster.id]] <- svd(cluster.mtx)$d^2/sum(svd(cluster.mtx)$d^2)
  rn.svd.vec.list[[cluster.id]] <-svd(cluster.mtx)$v
}

#Recover variance results and reformat
rn.svd.var.df <- plyr::ldply(rn.svd.var.list, function(x) x)
names(rn.svd.var.df) <- c("Cluster", paste("SVD", 1:(ncol(rn.svd.var.df) - 1), sep = ""))
rn.svd.var.df <- rn.svd.var.df %>% 
  gather(key = "SVD", value = "Variance", -Cluster) %>% 
  mutate(Variance = Variance * 100) 

#Recover singular vectors and reformat
rn.svd.vec.df <- plyr::ldply(rn.svd.vec.list, function(x) x)
names(rn.svd.vec.df) <- c("Cluster", paste("SVD", 1:(ncol(rn.svd.vec.df) - 1), sep = ""))
rn.svd.vec.df$SampleID <- colnames(rn.mtx)
rn.svd.vec.df <- rn.svd.vec.df %>% 
  gather(key = "SVD", value = "Score", -SampleID, -Cluster) %>% 
  inner_join(rn.map, by = "SampleID") %>% 
  inner_join(rn.svd.var.df, by = c("Cluster", "SVD"))
```

Reformat and explore correlations
```{r}
#Reformat RNA SVD results
rn.svd.tidy <- rn.svd.vec.df %>% 
  filter(SVD %in% paste("SVD", 1:1, sep = "")) %>% 
  dplyr::rename("RnaSampleID" = "SampleID") %>% 
  dplyr::select(RnaSampleID, Cluster, SVD, Score, Soil, Container) %>% 
  dplyr::rename("RnaCluster" = "Cluster",
                "RnaScore" = "Score",
                "RnaSVD" = "SVD")

#Reformat 16 SVD results
mb.svd.tidy <- mb.svd.vec.df %>%
  filter(SVD %in% paste("SVD", 1:1, sep = "")) %>% 
  dplyr::rename("MbSampleID" = "SampleID") %>% 
  dplyr::select(MbSampleID, Cluster, SVD, Score, Container) %>% 
  dplyr::rename("MbCluster" = "Cluster",
                "MbScore" = "Score",
                "MbSVD" = "SVD")

#merge
svd.master <- rn.svd.tidy %>% inner_join(mb.svd.tidy, by = "Container") %>% 
  mutate(Group = paste(RnaCluster, MbCluster, RnaSVD, MbSVD)) 

#Plot correlations
svd.master %>% 
  ggplot(aes(RnaScore, MbScore)) +
  geom_point(aes(shape = Soil, fill = Soil)) +
  geom_smooth(method = "lm", color = "black") +
  facet_wrap(~Group) +
  scale_fill_brewer(palette = "Set2",
                    guide = guide_legend(title.hjust = 0.5, title.position = "top", ncol = 1)) +
  scale_shape_manual(name = "", values = c(21,22,24),
                     guide = guide_legend(title.hjust = 0.5, title.position = "top", nrow = 1)) 


svd.master %>% 
  ggplot(aes(RnaScore, MbScore)) +
  geom_point(aes(shape = Soil, fill = Soil)) +
  geom_smooth(method = "lm", color = "black") +
  facet_wrap(~Group) +
  scale_fill_brewer(palette = "Set2",
                    guide = guide_legend(title.hjust = 0.5, title.position = "top", ncol = 1)) +
  scale_shape_manual(name = "", values = c(21,22,24),
                     guide = guide_legend(title.hjust = 0.5, title.position = "top", nrow = 1)) 
```

Test correlations
```{r}
run_cor_svd <- function(df){
  rbind(cor.test(df$RnaScore, df$MbScore, alternative = "two.sided", method = "pearson") %>% tidy())
}

supp.table.svd <- svd.master %>% 
  group_by(Group) %>% 
  nest() %>% 
  mutate(Correlation = map(data, run_cor_svd)) %>% 
  unnest(Correlation) %>% 
  ungroup() %>% 
  mutate(p.adj = p.adjust(p.value, method = "fdr")) %>% 
  #filter(p.adj < .05) %>% 
  #dplyr::select(Group, p.adj) %>% 
  arrange(p.adj) %>% 
  mutate(GeneCluster = str_sub(Group, 14,14),
         OTUCluster = str_sub(Group, 28,28)) %>% 
  dplyr::select(GeneCluster, OTUCluster, statistic, p.value, p.adj)

supp.table.svd

write.table(supp.table.svd, "../Tables/supp.gene.otu.svd.tsv", quote = F, sep = "\t", row.names = F, col.names = T)  
```

The two most significant correlations involved OTU Cluster 2. Generate figure with correlation patterns
```{r}
#Subset SVD results
mb.svd.var.1 <- filter(mb.svd.var.df, SVD == "SVD1")
rn.svd.var.1 <- filter(rn.svd.var.df, SVD == "SVD1")

#Plot correlation with gene Cluster 1
svd.p1 <- svd.master %>% 
  filter(Group =="Gene Cluster 1 OTU Cluster 2 SVD1 SVD1") %>% 
  ggplot(aes(MbScore, RnaScore)) +
  geom_point(aes(shape = Soil, fill = Soil)) +
  geom_smooth(method = "lm", color = "black", se = F) +
  scale_fill_brewer(name = "Soil", palette = "Set2",
                    guide = guide_legend(title.hjust = 0.5, title.position = "top", nrow = 1)) +
  scale_shape_manual(name = "Soil", values = c(21,22,24),
                     guide = guide_legend(title.hjust = 0.5, title.position = "top", nrow = 1)) +
  xlab(paste("OTU Cluster 2\nSV1 (", round(filter(mb.svd.var.1, Cluster == "OTU Cluster 2")$Variance, 1), "%)", sep = "")) +
  ylab(paste("Gene Cluster 1\nSV1 (", round(filter(rn.svd.var.1, Cluster == "Gene Cluster 1")$Variance, 1), "%)", sep = "")) +
  theme_bw() +
  theme(text = element_text(size = 11),
        legend.position = "none")

#Plot correlation with gene Cluster 1. Note that we switched the direction of the scores for visualization purposes so it fits the opposite trends displayed by the heatmap.
svd.p2 <- svd.master %>% 
  mutate(Soil = fct_recode(Soil,
                           "Arbuckle" = "Arb", 
                           "Biggs" = "Big",
                           "Davis" = "Dav")) %>% 
  mutate(RnaScore = -RnaScore) %>% 
  filter(Group =="Gene Cluster 2 OTU Cluster 2 SVD1 SVD1") %>% 
  ggplot(aes(MbScore, RnaScore)) +
  geom_point(aes(shape = Soil, fill = Soil)) +
  geom_smooth(method = "lm", color = "black", se = F) +
  scale_fill_brewer(name = "Soil", palette = "Set2",
                    guide = guide_legend(title.hjust = 0.5, title.position = "top", nrow = 3)) +
  scale_shape_manual(name = "Soil", values = c(21,22,24),
                     guide = guide_legend(title.hjust = 0.5, title.position = "top", nrow = 3)) +
  xlab(paste("OTU Cluster 2\nSV1 (", round(filter(mb.svd.var.1, Cluster == "OTU Cluster 2")$Variance, 1), "%)", sep = "")) +
  ylab(paste("Gene Cluster 2\nSV1 (", round(filter(rn.svd.var.1, Cluster == "Gene Cluster 2")$Variance, 1), "%)", sep = "")) +
  theme_bw() +
  theme(text = element_text(size = 11),
        legend.position = "none")


svd.p <- plot_grid(get_legend(svd.p2 + theme(legend.position = "top", text = element_text(size = 13))), svd.p1, svd.p2, 
                   nrow = 3, rel_heights = c(2, 3.5, 3.5), labels = c(NA, "(c)", NA), label_size = 15)
svd.p
```

```{r}
svd.cor.tidy <- svd.master %>% 
  group_by(Group) %>% 
  nest() %>% 
  mutate(Correlation = map(data, run_cor_svd)) %>% 
  unnest(Correlation) %>% 
  ungroup() %>% 
  mutate(p.adj = p.adjust(p.value, method = "fdr")) %>% 
  dplyr::select(Group, statistic, p.adj) %>% 
  mutate(statistic = round(statistic, 2)) %>% 
  mutate(p.adj = round(p.adj, 4)) %>% 
  mutate(RnaCluster = str_extract(Group, "Gene Cluster \\d")) %>% 
  mutate(MbCluster = str_extract(Group, "OTU Cluster \\d")) %>% 
  inner_join(dplyr::select(mb.svd.var.1, Cluster, Variance), by = c("MbCluster" = "Cluster")) %>% 
  dplyr::rename("MbVariance" = "Variance") %>% 
  inner_join(dplyr::select(rn.svd.var.1, Cluster, Variance), by = c("RnaCluster" = "Cluster")) %>% 
  dplyr::rename("RnaVariance" = "Variance") %>% 
  mutate(MbClusterFull = paste(MbCluster, " (", round(MbVariance, 1), "%)", sep = "")) %>% 
  mutate(RnaClusterFull = paste(RnaCluster, " (", round(RnaVariance, 1), "%)", sep = "")) %>% 
  mutate(StatsFull = paste("t = ", statistic, ", P = ", p.adj, sep = "")) 

#476:722
svd.master %>% 
  mutate(RnaScore = ifelse(RnaCluster == "Gene Cluster 2", -RnaScore, RnaScore)) %>% 
  inner_join(svd.cor.tidy, by = "Group") %>% 
  mutate(Soil = fct_recode(Soil,
                          "Arbuckle" = "Arb",
                          "Biggs" = "Big",
                          "Davis" = "Dav")) %>% 
  # inner_join(dplyr::select(mb.svd.var.1, Cluster, Variance), by = c("MbCluster" = "Cluster")) %>% 
  # dplyr::rename("MbVariance" = "Variance") %>% 
  # inner_join(dplyr::select(rn.svd.var.1, Cluster, Variance), by = c("RnaCluster" = "Cluster")) %>% 
  # dplyr::rename("RnaVariance" = "Variance") %>% 
  # mutate(MbClusterFull = paste(MbCluster, " (", round(MbVariance, 1), "%)", sep = "")) %>% 
  # mutate(RnaClusterFull = paste(RnaCluster, " (", round(RnaVariance, 1), "%)", sep = "")) %>% 
  # mutate(StatsFull = paste("t = ", statistic, ", P = ", p.adj, sep = "")) %>% 
  ggplot(aes(RnaScore, MbScore)) +
  geom_point(aes(shape = Soil, fill = Soil)) +
  geom_smooth(method = "lm", color = "black") +
  geom_text(data = svd.cor.tidy, aes(-.035, 0.5, label = StatsFull), size = 3) +
  xlab("Gene Cluster SVD 1") + 
  ylab("OTU Cluster SVD 1") +
  #facet_wrap(~Group) +
  facet_grid(MbClusterFull ~ RnaClusterFull) +
  scale_fill_brewer(palette = "Set2") +
  scale_shape_manual(name = "Soil", values = c(21,22,24))  +
  guides(fill = FALSE) +
  theme_bw() +
  theme(text = element_text(size = 12),
        legend.position = "top",
        strip.text = element_text(size = 9, colour = "white"),
        strip.background = element_rect(fill="gray25"))
```

Perform correlation analyses between OTU-gene pairs
```{r}
#Generate mapping file linking 16S and RNASeq samples
pairs.map <- mb.map %>% 
  dplyr::rename("MbSampleID" = "SampleID") %>% 
  inner_join(dplyr::select(rn.map, SampleID, Container), by = "Container") %>% 
  dplyr::rename("RnSampleID" = "SampleID") %>% 
  group_by(MbSampleID, RnSampleID, Container, Soil) %>% 
  dplyr::count() %>% 
  ungroup() %>% 
  dplyr::select(-n)

#Reformat z-transformed OTU matrix
mb.tidy <- mb.mtx %>% 
  tidy_otu() %>% 
  dplyr::rename("Abundance" = "Count") %>% 
  dplyr::rename("MbSampleID" = "SampleID")

#Reformat z-transformed gene matrix
rn.tidy <- rn.mtx %>% 
  tidy_otu() %>% 
  dplyr::rename("Expression" = "Count") %>% 
  dplyr::rename("RnSampleID" = "SampleID") %>% 
  dplyr::rename("gene" = "OTU_ID")

#Merge z-transformed OTU and gene data
pairs.tidy <- mb.tidy %>% 
  inner_join(pairs.map, by = "MbSampleID") %>% 
  inner_join(rn.tidy, by ="RnSampleID") %>% 
  left_join(dplyr::select(mb.clust, OTU_ID, Cluster, Core)) %>% 
  dplyr::rename("MbCluster" = "Cluster") %>% 
  left_join(dplyr::select(rn.clust, gene, Cluster, family, Symbol)) %>% 
  dplyr::rename("RnCluster" = "Cluster")

#Filter pairs that only include OTUs from Cluster 2
pairs.filt <- pairs.tidy %>% 
  filter(MbCluster == "OTU Cluster 2") 

#Run correlations
run_cor_pairs <- function(df){
  rbind(cor.test(df$Expression, df$Abundance, alternative = "two.sided", method = "pearson") %>% tidy())
}

pairs.res <- pairs.filt %>% 
  group_by(MbCluster, OTU_ID, RnCluster, gene) %>% 
  nest() %>% 
  mutate(Correlation = map(data, run_cor_pairs)) %>% 
  unnest(Correlation) %>% 
  ungroup() %>% 
  mutate(p.adj = p.adjust(p.value, method = "fdr"))

#How many significant correlations involved a core OTU
pairs.res %>% 
  filter(p.adj < 0.05) %>% 
  mutate(Core = OTU_ID %in% core.ids) %>% 
  group_by(Core) %>% 
  dplyr::count()
```

Explore OTUs with significant correlations
```{r}
#How many OTUs
pairs.res %>% 
  filter(p.adj < 0.05) %>% 
  group_by(OTU_ID) %>% 
  dplyr::count() %>% 
  arrange(-n)

#Filter the top 10 with the most correlations
pairs.sig <- pairs.res %>% 
  filter(p.adj < 0.05) %>% 
  group_by(OTU_ID, MbCluster, RnCluster) %>% 
  summarise(nCluster = n()) %>% 
  group_by(OTU_ID) %>% 
  mutate(Total = sum(nCluster)) %>% 
  ungroup() %>% 
  mutate(Rank = rank(-Total)) %>% 
  arrange(-Total) %>% 
  filter(Rank < 21) %>% 
  inner_join(tax, by = "OTU_ID")

#How many phyla are covered by the top 10 OTUs
pairs.sig %>% 
  group_by(Phylum) %>% 
  dplyr::count()

pairs.sig %>% 
  group_by(OTU_ID, Phylum,Rank) %>%
  dplyr::count() %>% 
  arrange(-Rank)
  
pairs.alt <- pairs.sig %>% 
  group_by(OTU_ID, Rank) %>% 
  count() %>% 
  arrange(Rank) %>% 
  ungroup() %>% 
  dplyr::select(OTU_ID)
pairs.alt$OTU_ID2 <- LETTERS[1:10]

#Plot the top 10 OTUs
pairs.cor.p <- pairs.sig %>% 
  inner_join(pairs.alt, by = "OTU_ID") %>% 
  mutate(Core = OTU_ID %in% core.ids) %>% 
  #mutate(Taxonomy = paste(str_sub(Phylum,1,1), ":", Family, OTU_ID2)) %>% 
  mutate(Taxonomy = paste(Family, OTU_ID2)) %>% 
  #mutate(Taxonomy = paste(str_sub(Phylum,1,1), ":", Genus, "OTU")) %>% 
  # mutate(Taxonomy = ifelse(Core,
  #                          paste("*", OTU_ID, sep = " "),OTU_ID)) %>% 
  ggplot(aes(reorder(Taxonomy, Rank), nCluster, fill = RnCluster)) +
  geom_bar(stat = "identity") +
  geom_point(aes(y = 0, color = Phylum), size = 3) +
  #scale_fill_manual(name = "", values = brewer.pal(11, "Reds")[c(7,4)]) +
  scale_fill_manual(name = "", values = brewer.pal(11, "Set2")[c(4,6)]) +
  scale_color_manual(name = "", values = rcartocolor::carto_pal(8, "Bold")[c(2,3,7,8)]) +
  guides(color = F) +
  guides(fill = guide_legend(nrow = 2)) +
  #coord_flip() +
  xlab("OTU") +
  ylab("Number of significant correlations\nwith a differentially expressed gene") +
  theme_bw() +
  theme(text = element_text(size = 10),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  #theme(legend.position = c(0.75,0.25))
  theme(legend.position = "bottom")

pairs.cor.p
```

Explore LRRs that were part of significant correlations
```{r}
#Subset correlation results and rank LRRs based on how many significant correlations they were part of
pairs.ranks <- pairs.res %>% 
  filter(p.adj < 0.05) %>% 
  inner_join(rcptr, by = "gene") %>% 
  group_by(gene, family) %>% 
  summarise(Total = n()) %>% 
  group_by(family) %>% 
  mutate(Rank = rank(-Total, ties.method = "first")) %>% 
  ungroup() %>% 
  dplyr::select(gene, Rank)


#Plot top 10 LRRs highlighting how many correlations involved a coreOTU
lrr.p <- pairs.res %>% 
  filter(p.adj < 0.05) %>% 
  inner_join(rcptr, by = "gene") %>% 
  group_by(gene) %>% 
  mutate(Total = n()) %>% 
  inner_join(pairs.ranks, by = "gene") %>% 
  filter(Rank < 11) %>% 
  mutate(Core = ifelse(OTU_ID %in% core.ids, "Core OTU", "Non-core OTU")) %>% 
  mutate(family = fct_relevel(family, "LRR-RLK")) %>% 
  ggplot(aes(reorder(gene, Total), fill = Core)) +
  geom_bar() +
  facet_grid(. ~ family, scales = "free", space = "free") +
  scale_fill_manual(name = "", values = rcartocolor::carto_pal(8, "Bold")[c(3,6)]) +
  guides(fill = guide_legend(nrow = 3)) +
  ylab("Number of significant correlations\nwith a differentially abundant OTU") +
  xlab("Gene") +
  theme_bw() +
  theme(text = element_text(size = 11),
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(size = 9, colour = "white"),
        strip.background = element_rect(fill="gray25"),
        axis.ticks = element_blank()) +
  theme(legend.position = "bottom")

lrr.p

#How many correlations in the top 10 LRRs involved a core TU
pairs.res %>% 
  filter(p.adj < 0.05) %>% 
  inner_join(rcptr, by = "gene") %>% 
  group_by(gene) %>% 
  mutate(Total = n()) %>% 
  inner_join(pairs.ranks, by = "gene") %>% 
  filter(Rank < 11) %>% 
  mutate(Core = OTU_ID %in% core.ids) %>% 
  group_by(Core) %>% 
  dplyr::count()

lrr.p <- pairs.res %>% 
  filter(p.adj < 0.05) %>% 
  inner_join(rcptr, by = "gene") %>% 
  group_by(gene) %>% 
  mutate(Total = n()) %>% 
  inner_join(pairs.ranks, by = "gene") %>% 
  filter(Rank < 11) %>% 
  mutate(Core = ifelse(OTU_ID %in% core.ids, "Core OTU", "Non-core OTU")) %>% 
  mutate(family = fct_relevel(family, "LRR-RLK")) %>% 
  inner_join(tax, by = "OTU_ID") %>% 
  ggplot(aes(reorder(gene, -Total), fill = Phylum)) +
  geom_bar() +
  geom_point(aes(y = 0, color = RnCluster), size = 3) +
  facet_grid(. ~ family, scales = "free", space = "free") +
  scale_fill_manual(name = "", values = rcartocolor::carto_pal(8, "Bold")) +
  scale_color_manual(name = "", values = brewer.pal(11, "Set2")[c(4,6)]) +
  guides(color = F) +
  ylab("Number of significant correlations\nwith a differentially abundant OTU") +
  xlab("Gene") +
  theme_bw() +
  theme(text = element_text(size = 10),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        strip.text = element_text(size = 7, colour = "white"),
        strip.background = element_rect(fill="gray25")) +
  theme(legend.position = "bottom")


lrr.p
```

```{r}
pairs.res %>% 
  filter(p.adj < 0.05) %>% 
  inner_join(rcptr, by = "gene") %>% 
  group_by(gene,family) %>% 
  dplyr::count() %>% 
  group_by(family) %>% 
  dplyr::count()

pairs.res %>% 
  filter(p.adj < 0.05) %>% 
  inner_join(rcptr, by = "gene") %>% 
  inner_join(tax, by = "OTU_ID") %>% 
  group_by(OTU_ID) %>% 
  count()

pairs.res %>% 
  filter(p.adj < 0.05) %>% 
  inner_join(rcptr, by = "gene") %>% 
  inner_join(tax, by = "OTU_ID") %>% 
  group_by(OTU_ID) %>% 
  count() %>% 
  inner_join(tax, by = "OTU_ID") %>% 
  group_by(Phylum) %>% 
  count() %>% 
  arrange(-n)

pairs.res %>% 
  filter(p.adj < 0.05) %>% 
  inner_join(rcptr, by = "gene") %>% 
  inner_join(tax, by = "OTU_ID") %>% 
  group_by(OTU_ID) %>% 
  count() %>% 
  mutate(Core = OTU_ID %in% core.ids) %>% 
  group_by(Core) %>% 
  count() %>% 
  arrange(-n)
 

pairs.res %>% 
  filter(p.adj < 0.05) %>% 
  inner_join(rcptr, by = "gene") %>% 
  group_by(gene) %>% 
  mutate(Total = n()) %>% 
  inner_join(pairs.ranks, by = "gene") %>% 
  filter(Rank < 11) %>% 
  group_by(family, gene, Rank, RnCluster) %>% 
  dplyr::count() %>% 
  arrange(family, -Rank)

ino.res <- read.table("../Tables/supp.deg.deseq.tsv", sep = "\t", header = T)

pairs.res %>% 
  filter(p.adj < 0.05) %>% 
  inner_join(rcptr, by = "gene") %>% 
  group_by(gene) %>% 
  dplyr::count() %>% 
  inner_join(ino.res, by = "gene") %>% 
  filter(Tissue == "Root")
  group_by(gene) %>% 
  mutate(Total = n())
  
lrr.ids <- pairs.res %>% 
  filter(p.adj < 0.05) %>% 
  inner_join(rcptr, by = "gene") %>% 
  group_by(gene) %>% 
  dplyr::count() %>% 
  .$gene 

lrr.ids
```


Put it all together. Make sure that the OTU and gene dendrograms from the other 2 notebooks are loaded.
```{r}
top <- plot_grid(rn.dendro.p, mb.dendro.p, svd.p, ncol = 3, labels = c("(a)", "(b)", NA), label_size = 15)

top <- plot_grid(rn.dendro.p, mb.dendro.p, soil.go, ncol = 3, labels = c("(a)", "(b)", "(c)"), label_size = 15, rel_widths = c(1,1,2))

right  <- plot_grid(pairs.cor.p, lrr.p, ncol = 2, align = "h", axis = "tb", rel_widths = c(2,3.5), labels = c("(d)", "(e)"), label_size = 15)

##733:983
##795:978
plot_grid(top, right, nrow = 2, rel_heights = c(1,1))
```

Generate supplementary table
```{r}
supp.table.cor <- pairs.res %>% 
  filter(p.adj < 0.05) %>% 
  dplyr::select(-data, -parameter, -conf.low, -conf.high, -method, -alternative) %>% 
  dplyr::rename("OTU_Cluster" = "MbCluster",
                "Gene_Cluster" = "RnCluster") %>% 
  group_by(OTU_ID) %>% 
  mutate(Total = n()) %>% 
  arrange(-Total) %>% 
  dplyr::select(-Total) %>% 
  left_join(pairs.alt, by = "OTU_ID") %>% 
  dplyr::rename("Fig6d_ID" = "OTU_ID2") %>% 
  left_join(rcptr, by = "gene") %>% 
  dplyr::rename("ReceptorFamily" = "family") %>% 
  dplyr::select(-Symbol) %>% 
  mutate(CoreOTU = OTU_ID %in% core.ids) %>% 
  inner_join(tax, by = "OTU_ID") %>% 
  dplyr::select(-Assignment)

write.table(supp.table.cor, "../Tables/supp.gene.otu.cor.tsv", quote = F, sep = "\t", row.names = F, col.names = T)  
```

