Load libraries
```{r}
library(DESeq2)
library(biobroom)
library(ggdendro)
library(vegan)
library(RColorBrewer)
library(cowplot)
library(tidyverse)

source("../../General/rmb_functions.R")
source("../../General/parameters.R")
```

Load data
```{r}
map <- readRDS("../Data/GNOTO_MB_map.RDS") %>% 
  filter(Treatment != "NC") %>% 
  mutate(Set = case_when(Compartment == "BS" ~ "BS",
                         Compartment == "ES" & Treatment == "PC" ~ "ES.PC",
                         Compartment == "ES" & Treatment == "Ino" ~ "ES.Ino"))

otu <- readRDS("../Data/SILVAcr_otu.RDS")
tax <- readRDS("../Data/SILVAcr_tax.RDS")

# Remove low prevalent OTUs (occupancy < 5% samples)
otu <- otu[,colnames(otu) %in% map$SampleID]
otu <- otu[rowSums(otu) > 0.05*ncol(otu),]
```

Get relative abundances and top phyla
```{r}
# Calculate relative abundances
otu.ra <- rel_ab(otu)
otu.ra.tidy <- tidy_otu(otu.ra) %>% mutate(Count = Count/100)

# Get top 11 phyla
tax <- expand_proteo(tax)
top.tax <- get_top_taxa(otu.ra, tax, rank = "PhyClass", n = 11)
tax <- collapse_other(tax, top.tax)

# Pool all other taxa into a "low abundance" category
tax <- tax %>% 
  mutate(PhyClass2 = fct_recode(PhyClass2, "Low abundance" = "other")) %>% 
  mutate(PhyClass2 = fct_relevel(PhyClass2, "Low abundance"))
```

Calculate Bray-Curtis dissimilarities
```{r}
dist <- beta_div_dist(otu.ra %>% log_norm)
```
Perform hierarchical clustering in each subset of sample types
```{r}
#Define clustering algorithm
clust.method <- "average"

#Subset mapping file and distance matrix
map.bs <- filter(map, Set == "BS")
map.ei <- filter(map, Set == "ES.Ino")
map.es <- filter(map, Set == "ES.PC")

dist.bs <- dist[match(map.bs$SampleID, row.names(dist)), match(map.bs$SampleID, colnames(dist))]
dist.ei <- dist[match(map.ei$SampleID, row.names(dist)), match(map.ei$SampleID, colnames(dist))]
dist.es <- dist[match(map.es$SampleID, row.names(dist)), match(map.es$SampleID, colnames(dist))]

#Perform hierarchical clustering and extract dendrogram data for plotting
ddata.bs <- as.dendrogram(hclust(as.dist(dist.bs), method = clust.method)) %>% dendro_data()
labs.bs <- label(ddata.bs) 

ddata.ei <- as.dendrogram(hclust(as.dist(dist.ei), method = clust.method)) %>% dendro_data()
labs.ei <- label(ddata.ei) 

ddata.es <- as.dendrogram(hclust(as.dist(dist.es), method = clust.method)) %>% dendro_data()
labs.es <- label(ddata.es) 

# Merge data together and format
segments.all <- rbind(segment(ddata.bs) %>% mutate(Set = "BS"),
                      segment(ddata.ei) %>% mutate(Set = "ES.Ino"),
                      segment(ddata.es) %>% mutate(Set = "ES.PC")) %>% 
  mutate(Set = fct_recode(Set,
                          "Bulk soil" = "BS",
                          "Endosphere SI" = "ES.Ino",
                          "Endosphere SG" = "ES.PC")) %>%
  mutate(Set = fct_relevel(Set, "Bulk soil", "Endosphere SG", "Endosphere SI"))

labs.all <- rbind(labs.bs,
                  labs.ei,
                  labs.es) %>% 
  dplyr::rename("SampleID" = "label") %>% 
  inner_join(map, by = "SampleID") %>% 
  mutate(Soil = fct_recode(Soil,
                           Arbuckle = "Arb",
                           Biggs = "Big",
                           Davis = "Dav")) %>% 
  mutate(Set = fct_recode(Set,
                          "Bulk soil" = "BS",
                          "Endosphere SI" = "ES.Ino",
                          "Endosphere SG" = "ES.PC")) %>%
  mutate(Set = fct_relevel(Set, "Bulk soil", "Endosphere SG", "Endosphere SI"))


#Plot
dendro.soil <- segments.all %>% 
  ggplot() +
  geom_segment(aes(x=x, y=y, xend=xend, yend=yend)) +
  geom_point(data=labs.all,
             aes(x=x, y=0.05, shape = Soil, fill = Soil), size = 2, color = "black") +
  scale_fill_brewer(palette = "Set2",
                    guide = guide_legend(title.hjust = 0.5,title.position = "top")) +
  scale_shape_manual(values = c(21,22,24)) +
  theme_bw() +
  theme(text = element_text(size = 11),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank(),
        strip.text = element_text(size = 10, colour = "white"),
        strip.background = element_rect(fill="gray25"),
        legend.position = "left") +
  facet_grid(. ~ Set, scales = "free", space = "free")

dendro.soil
```

Identify prevalent vOTUs and perform core microbiome analysis
```{r}
#Generate data frame with occupancy and abundance data in soil-inoculated endospheres
occupancy.df <- otu.ra.tidy %>% 
  inner_join(map.ei, by = "SampleID") %>% 
  group_by(Soil, OTU_ID) %>% 
  summarise(Occupancy = sum(Count>0),
            MeanAb = mean(Count)) %>% 
  filter(Occupancy > 0)

#Determine the core microbiome by identifying prevalent OTUs (occupancy > 6)
core.df <- occupancy.df %>% 
  filter(Occupancy > 6) %>% 
  group_by(OTU_ID) %>% 
  mutate(nSoil = n()) %>% 
  mutate(Core = nSoil == 3)

core.ids <- core.df %>% filter(Core) %>% .$OTU_ID %>% unique()
saveRDS(core.ids, "../Data/core_otu_ids.RDS")

#Plot occupancy 
occupancy.soil <- occupancy.df %>%
  mutate(Core = OTU_ID %in% filter(core.df, Core)$OTU_ID) %>% 
  ggplot(aes(MeanAb, Occupancy, color = Soil)) +
  geom_point(alpha = 0.2) +
  geom_hline(yintercept = 6.5, linetype = 2) +
  scale_x_log10() +
  scale_color_brewer(palette = "Set2") +
  facet_grid(. ~ Soil) +
  xlab("Mean relative abundance in ES SI") +
  ylab("Occupancy in ES SI\n(# samples)") +
  scale_x_log10(labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  theme_bw() +
  theme(text = element_text(size = 11),
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(size = 10, colour = "white"),
        strip.background = element_rect(fill="gray25"),
        legend.position = "none")

#Plot Venn diagram 
venn.soil <- core.df.venn <- plot(eulerr::venn(list(Arbuckle = filter(core.df, Soil == "Arbuckle")$OTU_ID,
                Biggs = filter(core.df, Soil == "Biggs")$OTU_ID,
                Davis = filter(core.df, Soil == "Davis")$OTU_ID)),
     fills = c("white", "white", "white"),
     edges = RColorBrewer::brewer.pal(3, "Set2"), lex = 3,
     labels = list(fontfamily = "Helvetica",
                   cex = 0),
     quantities = list(fontfamily = "Helvetica",
                   cex = 0.8))

#Plot aggregated abundances of core microbiome across soil sources in soil-inoculated endospheres
ab.soil <- otu.ra.tidy %>% 
  inner_join(map.ei, by = "SampleID") %>% 
  group_by(Soil, OTU_ID) %>% 
  summarise(AggRelAb = mean(Count)) %>% 
  mutate(Core = OTU_ID %in% filter(core.df, Core)$OTU_ID) %>% 
  filter(OTU_ID %in% filter(core.df, Core)$OTU_ID) %>% 
  group_by(Soil) %>% 
  summarise(TotalRelAb = sum(AggRelAb)) %>% 
  ggplot(aes(Soil, TotalRelAb, fill = Soil)) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette = "Set2") +
  ylab("Relative abundance\nof core OTUs in ES SI") +
  theme_bw() +
  theme(text = element_text(size = 11),
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank(),
        legend.position = "none")
```
Put it all together
```{r}
bottom <- plot_grid(occupancy.soil, NA, venn.soil, NA, ab.soil, nrow = 1, rel_widths = c(6,0.5,2,0.5,2.5), align = "h", axis = "b", labels = c("(b)",NA, "(c)",NA,"(d)"), label_size = 15)

#709:408
plot_grid(dendro.soil, bottom, nrow = 2, labels = c("(a)", NA), label_size = 15, rel_heights = c(2,3))
```


Generate a data frame with the abundance ranks of each OTU across soil types and growth conditions
```{r}
rank.diff <- otu.ra.tidy %>% 
  group_by(SampleID) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(Compartment == "ES") %>% 
  group_by(Treatment, Soil, OTU_ID) %>% 
  summarise(MeanRelAb = mean(Count)) %>% 
  filter(MeanRelAb > 0) %>% 
  group_by(Treatment, Soil) %>% 
  mutate(Rank = rank(-MeanRelAb, ties.method = "first")) %>% 
  select(-MeanRelAb) %>% 
  spread(key = Treatment, value = Rank) %>% 
  mutate(RankDiff = PC - Ino)
```

Plot rank abundance analysis for soil-grown endospheres, highlighting the overlap with soil-inoculated endospheres
```{r}
#Generate data frame with rank abundance data in SG ES 
rank.pc <- otu.ra.tidy %>% 
  group_by(SampleID) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(Compartment == "ES") %>% 
  group_by(Treatment, Soil, OTU_ID) %>% 
  summarise(MeanRelAb = mean(Count)) %>% 
  filter(MeanRelAb > 0) %>% 
  group_by(Treatment, Soil) %>% 
  mutate(Rank = rank(-MeanRelAb, ties.method = "first")) %>% 
  filter(Treatment == "PC") %>% 
  inner_join(select(rank.diff, Soil, OTU_ID, RankDiff), by = c("Soil", "OTU_ID")) %>% 
  mutate(Control = !is.na(RankDiff)) %>% 
  arrange(Soil, Rank) %>% 
  group_by(Soil) %>% 
  mutate(CumAb = cumsum(MeanRelAb))

#Define the rank threshold at which 0.9 total abundance is covered
rank.pc.thr <- rank.pc  %>% 
  filter(CumAb > 0.9) %>% 
  mutate(minRank = min(Rank)) %>% 
  filter(Rank == minRank)

#Calculate the fraction of OTUs also detected in SI ES considering all OTUs
rank.pc.ratio.all <- rank.pc %>% 
  group_by(Soil, Control) %>% 
  dplyr::count() %>% 
  group_by(Soil) %>% 
  mutate(Total = sum(n)) %>% 
  ungroup() %>% 
  filter(Control) %>% 
  mutate(AllRatio = paste(n, Total, sep = " / "))

#Calculate the fraction of OTUs also detected in SI ES considering only most abundant OTUs
rank.pc.ratio.top <- rank.pc %>% 
  filter(CumAb <= 0.9) %>% 
  group_by(Soil, Control) %>% 
  dplyr::count() %>% 
  group_by(Soil) %>% 
  mutate(Total = sum(n)) %>% 
  ungroup() %>% 
  filter(Control) %>% 
  mutate(TopRatio = paste(n, Total, sep = " / "))

#Merge
rank.pc.ratio <- rank.pc.ratio.all %>% 
  select(Soil, AllRatio) %>% 
  inner_join(select(rank.pc.ratio.top, Soil, TopRatio)) %>% 
  mutate(Ratio = paste("All OTUs: ", AllRatio, "\nTop OTUs: ", TopRatio, sep = ""))

#Plot rank abundance curve
rank.pc.p <- rank.pc %>% 
  ggplot() +
  geom_point(aes(Rank, MeanRelAb, color = Control), alpha = 0.5, size = 1) +
  geom_vline(data = rank.pc.thr, aes(xintercept = Rank), linetype = 2) +
  geom_text(data = rank.pc.ratio, aes(x = 500, y = 0.05, label = Ratio), size = 3, hjust = 0) +
  scale_y_log10(labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  scale_color_manual(name = "Detected\nin ES SI", values = c("gray50", "#FDBF6F")) +
  xlab("Abundance rank in ES SG") +
  ylab("Mean relative abundance\nin ES SG") +
  facet_grid(. ~ Soil) +
  theme_bw() +
  theme(text = element_text(size = 11),
        strip.text = element_blank(),
        legend.position = "right") 
  
#Plot histogram highlighting overlap with SI ES across different ranks
rank.pc.hist <- rank.pc %>% 
  mutate(Control = !is.na(RankDiff)) %>% 
  ggplot(aes(Rank, fill = Control)) +
  geom_histogram(binwidth = 20, ) +
  geom_vline(data = rank.pc.thr, aes(xintercept = Rank), linetype = 2) +
  scale_fill_manual(name = "Detected\nin ES SI", values = c("gray50", "#FDBF6F")) +
  facet_grid(. ~ Soil) +
  ylab("# of OTUs") +
  theme_bw() +
  theme(text = element_text(size = 11),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_text(size = 10, colour = "white"),
        strip.background = element_rect(fill="gray25"),
        legend.position = "none")


plot_grid(rank.pc.hist, rank.pc.p, nrow =2, align = "v", axis = "lr", rel_heights = c(1,2))
```

Plot rank abundance analysis for soil-inoculated endospheres, highlighting the overlap with soil-grown endospheres
```{r}
#Generate data frame with rank abundance data in SI ES
rank.ino <- otu.ra.tidy %>% 
  group_by(SampleID) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(Compartment == "ES") %>% 
  group_by(Treatment, Soil, OTU_ID) %>% 
  summarise(MeanRelAb = mean(Count)) %>% 
  filter(MeanRelAb > 0) %>% 
  group_by(Treatment, Soil) %>% 
  mutate(Rank = rank(-MeanRelAb, ties.method = "first")) %>% 
  filter(Treatment == "Ino") %>% 
  inner_join(select(rank.diff, Soil, OTU_ID, RankDiff), by = c("Soil", "OTU_ID")) %>% 
  mutate(Control = !is.na(RankDiff)) %>% 
  arrange(Soil, Rank) %>% 
  group_by(Soil) %>% 
  mutate(CumAb = cumsum(MeanRelAb))

#Define the rank threshold at which 0.9 total abundance is covered
rank.ino.thr <- rank.ino  %>% 
  filter(CumAb > 0.9) %>% 
  mutate(minRank = min(Rank)) %>% 
  filter(Rank == minRank)

#Calculate the fraction of OTUs also detected in SG ES considering all OTUs
rank.ino.ratio.all <- rank.ino %>% 
  group_by(Soil, Control) %>% 
  dplyr::count() %>% 
  group_by(Soil) %>% 
  mutate(Total = sum(n)) %>% 
  ungroup() %>% 
  filter(Control) %>% 
  mutate(AllRatio = paste(n, Total, sep = " / "))

#Calculate the fraction of OTUs also detected in SG ES considering only most abundant OTUs
rank.ino.ratio.top <- rank.ino %>% 
  filter(CumAb <= 0.9) %>% 
  group_by(Soil, Control) %>% 
  dplyr::count() %>% 
  group_by(Soil) %>% 
  mutate(Total = sum(n)) %>% 
  ungroup() %>% 
  filter(Control) %>% 
  mutate(TopRatio = paste(n, Total, sep = " / "))

#Merge
rank.ino.ratio <- rank.ino.ratio.all %>% 
  select(Soil, AllRatio) %>% 
  inner_join(select(rank.ino.ratio.top, Soil, TopRatio)) %>% 
  mutate(Ratio = paste("All OTUs: ", AllRatio, "\nTop OTUs: ", TopRatio, sep = ""))

#Plot rank abundance curve
rank.ino.p <- rank.ino %>% 
  ggplot() +
  geom_point(aes(Rank, MeanRelAb, color = Control), alpha = 0.5, size = 1) +
  geom_vline(data = rank.ino.thr, aes(xintercept = Rank), linetype = 2) +
  geom_text(data = rank.ino.ratio, aes(x = 500, y = 0.05, label = Ratio), size = 3, hjust = 0) +
  scale_y_log10(labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  scale_color_manual(name = "Detected\nin ES SG", values = c("gray50", "#FF7F00")) +
  xlab("Abundance rank in ES SI") +
  ylab("Mean relative abundance\nin ES SI") +
  facet_grid(. ~ Soil) +
  theme_bw() +
  theme(text = element_text(size = 11),
        strip.text = element_blank(),
        legend.position = "right")

#Plot histogram highlighting overlap with SI ES across different ranks
rank.ino.hist <- rank.ino %>% 
  mutate(Control = !is.na(RankDiff)) %>% 
  ggplot(aes(Rank, fill = Control)) +
  geom_histogram(binwidth = 20, ) +
  geom_vline(data = rank.ino.thr, aes(xintercept = Rank), linetype = 2) +
  scale_fill_manual(name = "Detected\nin ES SG", values = c("gray50", "#FF7F00")) +
  facet_grid(. ~ Soil) +
  ylab("# of OTUs") +
  theme_bw() +
  theme(text = element_text(size = 11),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_text(size = 10, colour = "white"),
        strip.background = element_rect(fill="gray25"),
        legend.position = "none")

plot_grid(rank.ino.hist, rank.ino.p, nrow =2, align = "v", axis = "lr", rel_heights = c(1,2))
```

Perform differential abundance analysis testing differences between growth conditions in endosphere communities
```{r}
#Subset mapping file and OTU table
es.map <- map %>% 
  filter(Compartment == "ES") %>% 
  mutate(Group = paste(Treatment, Soil, sep = "."))

es.otu <- otu[,match(es.map$SampleID, colnames(otu))]
es.otu <- es.otu[rowSums(es.otu)>0,]


# Load into DESeq
es.dds <- DESeqDataSetFromMatrix(es.otu,
                                colData = es.map,
                                design = ~ Group)

es.dds <- DESeq(es.dds)

# Run contrasts
arb <- lfcShrink(es.dds, contrast = c("Group", "Ino.Arbuckle", "PC.Arbuckle")) %>% tidy() %>% mutate(Soil = "Arbuckle")
big <- lfcShrink(es.dds, contrast = c("Group", "Ino.Biggs", "PC.Biggs")) %>% tidy() %>% mutate(Soil = "Biggs")
dav <- lfcShrink(es.dds, contrast = c("Group", "Ino.Davis", "PC.Davis")) %>% tidy() %>% mutate(Soil = "Davis")

# Compile and perform adjustments after pooling all contrasts
es.res <- rbind(arb, big, dav) %>% 
  dplyr::rename("OTU_ID" = "gene") %>% 
  ungroup() %>% 
  filter(!is.na(p.adjusted)) %>% 
  mutate(p.adjusted2 = p.adjust(p.value, method = "fdr"))

#Generate supplementary table
supp.table <- es.res %>% 
  filter(p.adjusted2 < 0.05) %>% 
  mutate(Direction = ifelse(estimate > 0, "Enriched in SI", "Enriched in SG")) %>% 
  select(Soil, OTU_ID, Direction, everything()) %>% 
  select(-p.adjusted) %>% 
  dplyr::rename("p.adjusted" = "p.adjusted2") %>% 
  inner_join(tax, by = "OTU_ID") %>% 
  select(-Assignment, -PhyClass, -PhyClass2)

write.table(supp.table, "../Tables/supp.dao.si.vs.sg.tsv", quote = F, sep = "\t", row.names = F, col.names = T)
```

Plot differential abundance analysis results
```{r}
#Generate a data frame with the abundance correlations between SG and SI samples for the subset of OTUs shared between the two growth conditions
cor.df <- otu.ra.tidy %>% 
  group_by(SampleID) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(Compartment == "ES") %>% 
  group_by(Treatment, Soil, OTU_ID) %>% 
  summarise(MeanRelAb = mean(Count)) %>% 
  filter(MeanRelAb > 0) %>% 
  group_by(Treatment, Soil) %>% 
  mutate(Rank = rank(-MeanRelAb, ties.method = "first")) %>% 
  select(-Rank) %>% 
  spread(key = Treatment, value = MeanRelAb) %>% 
  filter(!is.na(Ino) & !is.na(PC)) %>% 
  left_join(es.res, by = c("Soil", "OTU_ID")) %>% 
  mutate(DAO = case_when(p.adjusted2 < 0.05 ~ TRUE,
                         p.adjusted2 >= 0.05 ~ FALSE,
                         is.na(p.adjusted2) ~ FALSE)) 

#Calculate the ratio of OTUs detected as differentially abundant 
cor.ratio <- cor.df %>% 
  group_by(Soil, DAO) %>% 
  dplyr::count() %>% 
  group_by(Soil) %>% 
  mutate(Total = sum(n)) %>% 
  filter(DAO) %>% 
  mutate(Ratio = paste(n, Total, sep = " / "))

#Plot
cor.p <- cor.df %>% 
  ggplot(aes(Ino, PC)) +
  geom_point(alpha = 0.5, aes(color = DAO), size = 1) +
  geom_abline(slope = 1, intercept = 0, linetype = 2) +
  geom_text(data = cor.ratio, aes(x = 0.00001, y = 0.1, label = Ratio), size = 3, hjust = 0) +
  scale_x_log10(labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  scale_y_log10(labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  xlab("Mean relative abundance in ES SI") +
  ylab("Mean relative abundance\nin ES SG") +
  facet_grid(. ~ Soil) +
  scale_color_manual(name = "Differentially\nabundant\nSI vs SG", values = c("gray50", "dodgerblue2")) +
  theme_bw() +
  theme(text = element_text(size = 11),
        strip.text = element_text(size = 10, colour = "white"),
        strip.background = element_rect(fill="gray25"))

cor.p
```

Put it all together
685:916
```{r}
plot_grid(rank.pc.hist, rank.pc.p, rank.ino.hist, rank.ino.p, cor.p, nrow =5, align = "v", axis = "lr", rel_heights = c(1,2,1,2,2), labels = c("(a)", NA, "(b)", NA, "(c)"), label_size = 15)
```