Load libraries
```{r}
library(cowplot)
library(tidyverse)
```

Load DESeq results and annotate identify which genes are part of the core response
```{r}
trt.res <- readRDS("../Data/msu_trt_res.RDS") %>% 
  filter(!is.na(p.adjusted)) %>% 
  filter(baseMean > 0) %>% 
  group_by(Tissue) %>% 
  mutate(p.adjusted2 = p.adjust(p.value, method = "fdr")) %>% 
  ungroup() %>% 
  mutate(Direction = case_when(p.adjusted2 >= 0.05 | is.na(p.adjusted2) ~ "NS",
                               p.adjusted2 < 0.05 & Tissue == "Leaf" & estimate > 0 ~ "L.Up",
                               p.adjusted2 < 0.05 & Tissue == "Leaf" & estimate < 0 ~ "L.Dn",
                               p.adjusted2 < 0.05 & Tissue == "Root" & estimate > 0 ~ "R.Up",
                               p.adjusted2 < 0.05 & Tissue == "Root" & estimate < 0 ~ "R.Dn")) %>% 
  mutate(Significant = p.adjusted2 < 0.05) %>% 
  group_by(Tissue, Direction, gene) %>% 
  mutate(Total = sum(Significant)) %>% 
  ungroup() %>% 
  mutate(Core = ifelse(Total == 3, TRUE, FALSE)) %>% 
  mutate(estimate = ifelse(abs(estimate) > 5, 5*sign(estimate), estimate)) %>% 
  mutate(Soil = fct_recode(Soil,
                           "Arbuckle" = "Arb",
                           "Biggs" = "Big",
                           "Davis" = "Dav")) %>% 
  mutate(Tissue = fct_relevel(Tissue, "Root")) 



```


Load LRR annotations and format
```{r}
# Load data from 
lrr <- read.table("../../General/lrr.tsv", header = T, sep = "\t", quote = "", comment.char = "")
nlr.1 <- filter(lrr, Gene.Fam == "NLR")$ID_MSU %>% unique() %>% as.character()

# Load data from
nlr.2 <- read.table("../../General/nlr.tsv", header = F)$V1 %>% as.character()
nlr.2 <- paste("LOC_", nlr.2, sep = "")

# Get all NLRs across the two annotation sources
nlr <- unique(c(nlr.1, nlr.2))
nlr %>% length()

# Get RLKs
rlk <- filter(lrr, Gene.Fam == "LRR-RLK")$ID_MSU %>% unique() %>% as.character()

# Load the Symbol names from the funrice database
funrice <- read.table("../../General/geneKeyword.table.txt", header = T, sep = "\t", quote = "")
funrice.ids <- select(funrice, MSU, Symbol) %>% group_by(MSU,Symbol) %>% count %>% ungroup() %>% select(-n)

# Compile 
rcptr <- rbind(data.frame(gene = nlr, family = "NLR"),
               data.frame(gene = rlk, family = "LRR-RLK")) %>% 
  left_join(funrice.ids, by = c("gene" = "MSU"))
```

Function to perform hypergeometric test
```{r}
gene.universe <- trt.res %>% filter(Tissue == "Root") %>% group_by(gene) %>% dplyr::count()

rcptr.cp <- rcptr %>% 
  right_join(gene.universe, by = "gene") %>% 
  select(gene, family) %>% 
  mutate(family = ifelse(is.na(family), "Unannotated", as.character(family))) %>% 
  mutate(family2 = family)

get_lrr <- function(x){
  genes <- x$gene
  res = clusterProfiler::enricher(gene = genes, 
             universe = gene.universe$gene, 
             TERM2GENE = select(rcptr.cp, family, gene), 
             TERM2NAME = select(rcptr.cp, family, family2),
             pvalueCutoff = 1,
             qvalueCutoff = 1,
             minGSSize = 10,
             maxGSSize = 5000)
  as.data.frame(res)
}
```

Generate a nested data frame and run hypergeometric tests within subsets
```{r}
trt.sig <- trt.res %>% 
  filter(p.adjusted2 < 0.05)

contra.df <- trt.sig %>% 
  mutate(Direction = ifelse(estimate > 0, "Upregulated", "Downregulated")) 

core.nest <- contra.df %>% 
  group_by(Tissue, Direction, Soil) %>% 
  nest() %>% 
  mutate(hyperG = map(data, get_lrr))

core.nest %>% 
  unnest(hyperG) %>% 
  group_by(Tissue) %>% 
  mutate(p.adjusted2 = p.adjust(pvalue, method = "fdr")) %>% 
  filter(p.adjusted2 < 0.05) %>% 
  group_by(Tissue, Direction) %>% 
  mutate(Rank = rank(pvalue)) %>% 
  ggplot(aes(reorder(ID, -Rank), -log(p.adjust))) +
  geom_bar(stat = "identity") +
  coord_flip() +
  facet_grid(Tissue + Direction + Soil ~ ., scales = "free_y")
```

Generate supplementary table with all DEGs annotated as LRR
```{r}
supp.table.lrr <- trt.sig %>% 
  filter(Tissue == "Root") %>% 
  mutate(Direction = ifelse(estimate > 0, "Upregulated by MB", "Downregulated by MB")) %>% 
  group_by(Tissue, Direction, gene) %>% 
  mutate(n = n()) %>% 
  ungroup() %>% 
  mutate(Core = n == 3) %>% 
  inner_join(rcptr, by = "gene") %>% 
  select(Tissue, Soil, family, gene, Symbol, Direction, Core, baseMean:p.value, p.adjusted2) %>% 
  rename("log2FoldChange_SIvsMI" = "estimate",
         "p.adjusted" = "p.adjusted2") 

write.table(supp.table.lrr, "../Tables/supp.lrr.tsv", quote = F, sep = "\t", row.names = F, col.names = T)
```
Check how many LRRs were found within each subset
```{r}
supp.table.lrr %>% 
  group_by(family, Core, Direction, gene) %>% 
  count() %>% 
  ungroup() %>% 
  select(-n) %>% 
  group_by(family, Direction, Core) %>% 
  count()
```

Check LRRs with a known name
```{r}
supp.table.lrr %>% 
  filter(!is.na(Symbol)) %>% 
  arrange(family, Core, Direction, Symbol) %>% 
  filter(Core & str_detect(Direction, "Down"))
```

Plot 
```{r}
# Histogram displaying effect size
lrr.hist.p <- trt.res %>% 
  inner_join(rcptr, by = "gene") %>% 
  filter(Significant) %>% 
  mutate(family = case_when(family == "NLR" ~ "NLR\n(n = 427)",
                            family == "LRR-RLK" ~ "LRR-RLK\n(n = 303)")) %>% 
  mutate(family = fct_relevel(family, "NLR\n(n = 427)")) %>% 
  mutate(Core = ifelse(Core, "Core", as.character(Direction))) %>% 
  mutate(Core = fct_relevel(Core, "R.Up", "R.Dn", "L.Up", "L.Dn", "Core")) %>% 
  filter(Tissue == "Root") %>% 
  ggplot(aes(estimate)) +
  geom_histogram(aes(fill = Direction, color = Core)) +
  facet_grid(family ~ Soil, , scales = "free") + 
  #scale_color_manual(values = c("white", "black")) +
  scale_fill_manual(values = c("#8C510A","#DFC27D","#4D9221","#B8E186","gray75"),
                     breaks = c("R.Up", "R.Dn", "L.Up", "L.Dn", "NS"),
                     guide = F) +
  scale_color_manual(values = c("#8C510A","#DFC27D","#4D9221","#B8E186","gray15"),
                     breaks = c("R.Up", "R.Dn", "L.Up", "L.Dn", "Core"),
                     labels = c("Root genes\nupreg. by MB",
                                "Root genes\ndownreg. by MB",
                                "Leaf genes\nupreg. by MB",
                                "Leaf genes\ndownreg. by MB",
                                "Core"),
                     name = "",
                     drop = F) +
  xlab("log2 fold change (SI vs MI)") +
  ylab("Frequency (# genes)") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(size = 13),
        strip.text = element_text(size = 10, colour = "white"),
        strip.background = element_rect(fill="gray25"),
        legend.position = "bottom")


lrr.hist.p

# Data frame for donut plots
lrr.total <- trt.res %>% 
  filter(Significant) %>% 
  inner_join(rcptr, by = "gene") %>% 
  filter(Tissue == "Root") %>% 
  group_by(Tissue, Soil, family) %>% 
  summarise(Total = n()) %>% 
  mutate(Soil = fct_recode(Soil,
                           "Arbuckle" = "Arb",
                           "Biggs" = "Big",
                           "Davis" = "Dav")) %>% 
  mutate(Tissue = fct_relevel(Tissue, "Root"))

# Donut plots
lrr.donut.p <- trt.res %>% 
  filter(Significant) %>% 
  inner_join(rcptr, by = "gene") %>% 
  filter(Tissue == "Root") %>% 
  group_by(Tissue, Soil, Direction, family) %>% 
  summarise(Count = n()) %>% 
  group_by(Tissue, Soil, family) %>% 
  mutate(Fraction = Count / sum(Count)) %>% 
  mutate(ymax = cumsum(Fraction),
         nPhy = n()) %>% 
  mutate(ymin = c(0, ymax[1:nPhy - 1])) %>% 
  mutate(ymid = ymin + (ymax -ymin)/2) %>% 
  mutate(Count2 = ifelse(Direction == "R.Dn", paste(Count,"*", sep = ""), Count)) %>% 
  ggplot() +
  geom_rect(aes(ymax=ymax, ymin=ymin, xmax= 4, xmin= 3, fill= Direction)) +
  geom_text(aes(x = 3.5, y = ymid, label = Count2, color = Direction)) +
  geom_text(data = lrr.total, aes(2, 0, label = Total), size = 5) +
  scale_fill_manual(values = c("#8C510A","#DFC27D","#4D9221","#B8E186","gray75"),
                     breaks = c("R.Up", "R.Dn", "L.Up", "L.Dn", "NS")) +
  scale_color_manual(values = c("black", "white")) +
  coord_polar(theta="y") + 
  xlim(c(2, 4)) +
  facet_wrap(. ~ family + Soil, nrow = 1) +
  theme_void() +
  theme(text = element_text(size = 15),
        legend.position = "none",
        legend.title = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank())

lrr.donut.p
```