Load libraries
```{r}
library(topGO)
library(cowplot)
library(tidytext)
library(tidyverse)
```

Load data
```{r}
map <- readRDS("../Data/RNASeq_map.RDS")

trt.res <- readRDS("../Data/msu_trt_res.RDS") %>% 
  filter(!is.na(p.adjusted)) %>% 
  filter(baseMean > 0) %>% 
  group_by(Tissue) %>% 
  mutate(p.adjusted2 = p.adjust(p.value, method = "fdr")) %>% 
  ungroup()

go.defs <- readRDS("../Data/go_defs.rds")
go.topgo <- readRDS("../Data/msu_topgo.RDS")

pfam <- readRDS("../Data/msu_pfam.RDS")
```

Create data frame with the list of genes in each subset of the core response
```{r}
trt.sig <- trt.res %>% 
  filter(p.adjusted2 < 0.05)

contra.df <- trt.sig %>% 
    mutate(Direction = ifelse(estimate > 0, "Upregulated", "Downregulated")) %>% 
  group_by(Tissue, Direction, gene) %>% 
  dplyr::count() %>% 
  filter(n == 3)
```

Create topGO objects
```{r}
assayed.genes <- trt.res$gene %>% unique()
tmp.sig.genes <- sample(assayed.genes, 100)
tmp.gene.vec <- as.integer(assayed.genes %in% tmp.sig.genes)
names(tmp.gene.vec) <- assayed.genes

# For biological process ontology
GOdata.BP <- new("topGOdata",
                description = "test",
                ontology = "BP",
                allGenes = as.factor(tmp.gene.vec),
                annot = annFUN.gene2GO,
                gene2GO = go.topgo,
                nodeSize = 1)

# For molecular function ontology
GOdata.MF <- new("topGOdata",
                description = "test",
                ontology = "MF",
                allGenes = as.factor(tmp.gene.vec),
                annot = annFUN.gene2GO,
                gene2GO = go.topgo,
                nodeSize = 1)
```

Make function to run topGO pipeline
```{r}
run_topgo <- function(df, GOdata){
  sig.genes <- df$gene
  gene.vec <- as.integer(assayed.genes %in% sig.genes)
  names(gene.vec) <- assayed.genes

  GOdata <- updateGenes(GOdata, as.factor(gene.vec))

  classic <- runTest(GOdata, algorithm = "classic", statistic = "fisher") 
  w01 <- runTest(GOdata, algorithm = "weight01", statistic = "fisher")
  
  GenTable(
    GOdata,
    classic = classic,
    w01 = w01,
    orderBy = "classic",
    ranksOf = "classic",
    topNodes = length(usedGO(GOdata))
    ) %>%
    mutate(classic = as.numeric(classic),
    w01 = as.numeric(w01)) %>% 
    rename(GO = GO.ID)
} 
```

Function to perform hypergeometric tests on Pfam terms
```{r}
gene.universe <- trt.res %>% group_by(gene) %>% dplyr::count()

pfam.clust <- pfam %>% 
  right_join(gene.universe, by = "gene") %>% 
  select(-n) %>% 
  mutate(pfam = ifelse(is.na(pfam), "Unannotated", as.character(pfam)),
         name = ifelse(is.na(name), "Unannotated", as.character(name))) 

get_pfam <- function(x){
  genes <- x$gene
  res = clusterProfiler::enricher(gene = genes, 
             universe = gene.universe$gene, 
             TERM2GENE = select(pfam.clust, pfam, gene), 
             TERM2NAME = select(pfam.clust, pfam, name),
             pvalueCutoff = 1,
             qvalueCutoff = 1,
             minGSSize = 10,
             maxGSSize = 5000)
  as.data.frame(res)
}
```

Generate a nested data frame to run enrichment analyses
```{r}
core.nest <- contra.df %>% 
  group_by(Tissue, Direction) %>% 
  nest() %>% 
  mutate(hyperG = map(data, get_pfam),
         GO_BP = map(data, run_topgo, GOdata.BP),
         GO_MF = map(data, run_topgo, GOdata.MF))
```

Unnest and compile GO results
```{r}
go.res.bp <- core.nest %>% 
  unnest(GO_BP) %>% 
  mutate(Ontology = "Biological Process")

go.res.mf <- core.nest %>% 
  unnest(GO_MF) %>% 
  mutate(Ontology = "Molecular Function")

go.res <- rbind(go.res.bp, go.res.mf)
```

Generate supplementary table with all significant GO terms
```{r}
supp.table.go <- go.res %>% 
  ungroup() %>% 
  group_by(Tissue, Ontology) %>% 
  mutate(w012 = p.adjust(w01, method = "fdr")) %>% 
  filter(w012 < 0.05) %>% 
  ungroup() %>% 
  select(Tissue, Direction, GO, Term, Significant, Annotated, w012, Ontology) %>% 
  rename("ID" = "GO",
         "Description" = "Term",
         "GeneSet" = "Significant",
         "GeneUniverse" = "Annotated",
         "p.val" = "w012",
         "Annotation" = "Ontology")

write.table(supp.table.go, "../Tables/supp.go.tsv", quote = F, sep = "\t", row.names = F, col.names = T)
```

Subset GO results to only include top 5 most significant terms in each subset
```{r}
go.res.tidy <- supp.table.go %>% 
  group_by(Tissue, Direction, Annotation) %>% 
  mutate(Rank = rank(p.val)) %>%
  filter(Rank < 6) %>% 
  ungroup() 
```

Generate supplementary table with all significant Pfam terms
```{r}
supp.table.pfam <- core.nest %>% 
  unnest(hyperG) %>% 
  group_by(Tissue) %>% 
  mutate(p.adjusted2 = p.adjust(pvalue, method = "fdr")) %>% 
  filter(p.adjusted2 < 0.05) %>% 
  separate(GeneRatio, c("GeneSet", "TotalSet"), sep = "/") %>% 
  separate(BgRatio, c("GeneUniverse", "TotalUniverse"), sep = "/") %>% 
  mutate(GeneSet = as.numeric(GeneSet),
         TotalSet = as.numeric(TotalSet),
         GeneUniverse = as.numeric(GeneUniverse),
         TotalUniverse = as.numeric(TotalUniverse)) %>% 
  select(Tissue, Direction, ID, Description, GeneSet, GeneUniverse, p.adjusted2) %>% 
  rename("p.val" = "p.adjusted2") %>% 
  mutate(Annotation = "Pfam")

write.table(supp.table.pfam, "../Tables/supp.pfam.tsv", quote = F, sep = "\t", row.names = F, col.names = T)
```

Subset Pfam results to only include top 5 most significant terms in each subset
```{r}
pfam.res.tidy <- supp.table.pfam %>% 
  group_by(Tissue, Direction) %>% 
  mutate(Rank = rank(p.val)) %>% 
  filter(Rank < 6) %>% 
  ungroup()
```

Compile results and plot
```{r}
all.res <- rbind(pfam.res.tidy %>% ungroup(), go.res.tidy %>% ungroup())

root.p <- all.res %>% 
  mutate(Color = interaction(Tissue, Direction)) %>% 
  mutate(Term = reorder_within(Description, -log10(p.val), Color)) %>% 
  mutate(Set = paste(Tissue, Direction, sep = ".")) %>% 
  mutate(Set = fct_recode(Set,
                          "Root genes\nupreg. by MB" = "Root.Upregulated",
                          "Root genes\ndownreg. by MB" = "Root.Downregulated",
                          "Leaf genes\nupreg. by MB" = "Leaf.Upregulated",
                          "Leaf genes\ndownreg. by MB" = "Leaf.Downregulated")) %>% 
  mutate(Set = fct_relevel(Set,
                           "Root genes\nupreg. by MB", "Root genes\ndownreg. by MB",
                           "Leaf genes\nupreg. by MB", "Leaf genes\ndownreg. by MB")) %>% 
  mutate(Annotation = fct_recode(Annotation,
                                 "BP" = "Biological Process",
                                 "MF" = "Molecular Function")) %>% 
  mutate(Annotation = fct_relevel(Annotation,
                                  "Pfam")) %>% 
  filter(Tissue == "Root") %>% 
  ggplot(aes(Term, -log10(p.val), fill = Set)) +
  geom_bar(stat = "identity", color = "black") +
  scale_x_reordered() +
  facet_grid(Direction + Annotation ~ Tissue, scales = "free", space = "free") +
  coord_flip() +
  scale_fill_manual(name = "",values = c("#8C510A", "#DFC27D", "#4D9221","#B8E186"),
                    limits = c("Root genes\nupreg. by MB", "Root genes\ndownreg. by MB", "Leaf genes\nupreg. by MB", "Leaf genes\ndownreg. by MB")) +
  scale_color_manual(values = c("black", "white"), guide = F) +
  coord_flip() +
  ylab("-log10 p-value") +
  xlab("") +
  ylim(0,10) +
  theme_bw() +
  theme(text = element_text(size = 10),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")
  

leaf.p <- all.res %>% 
  mutate(Color = interaction(Tissue, Direction)) %>% 
  mutate(Term = reorder_within(Description, -log10(p.val), Color)) %>% 
  mutate(Set = paste(Tissue, Direction, sep = ".")) %>% 
  mutate(Set = fct_recode(Set,
                          "Root genes\nupreg. by MB" = "Root.Upregulated",
                          "Root genes\ndownreg. by MB" = "Root.Downregulated",
                          "Leaf genes\nupreg. by MB" = "Leaf.Upregulated",
                          "Leaf genes\ndownreg. by MB" = "Leaf.Downregulated")) %>% 
  mutate(Set = fct_relevel(Set,
                           "Root genes\nupreg. by MB", "Root genes\ndownreg. by MB",
                           "Leaf genes\nupreg. by MB", "Leaf genes\ndownreg. by MB")) %>% 
  mutate(Annotation = fct_recode(Annotation,
                                 "BP" = "Biological Process",
                                 "MF" = "Molecular Function")) %>% 
  mutate(Annotation = fct_relevel(Annotation,
                                  "Pfam")) %>% 
  filter(Tissue == "Leaf") %>% 
  ggplot(aes(Term, -log10(p.val), fill = Set)) +
  geom_bar(stat = "identity", color = "black") +
  scale_x_reordered() +
  facet_grid(Direction + Annotation ~ Tissue, scales = "free", space = "free") +
  coord_flip() +
  scale_fill_manual(name = "",values = c("#8C510A", "#DFC27D", "#4D9221","#B8E186"),
                    limits = c("Root genes\nupreg. by MB", "Root genes\ndownreg. by MB", "Leaf genes\nupreg. by MB", "Leaf genes\ndownreg. by MB")) +
  scale_color_manual(values = c("black", "white"), guide = F) +
  coord_flip() +
  ylab("-log10 p-value") +
  xlab("") +
  ylim(0,10) +
  theme_bw() +
  theme(text = element_text(size = 10),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")

cowplot::plot_grid(root.p, leaf.p, align = "vh", axis = "lrbt")
```