Load libraries
```{r}
library(DESeq2)
library(tidyverse)
library(biobroom)
```

Load data
```{r}
txi <- readRDS("../Data/txi_msu.rds")
map <- readRDS("../Data/RNASeq_map.RDS")

map <- map[match(colnames(txi$abundance), map$SampleID), ]
```

Generate DESeq object to test the effect of inoculating a soil microibome on the plant transcriptomes
```{r}
dds <- DESeqDataSetFromTximport(txi,
                                colData = map,
                                design = ~ Group)
```

Perform contrasts comparing soil-inoculated vs mock-inoculated transcriptomes in each leaf and soil type
```{r}
dds <- DESeq(dds)


leaf.arb <- results(dds, contrast = c("Group", "Leaf.Ino.Arb", "Leaf.NC.Arb")) %>% tidy() %>% mutate(Tissue = "Leaf", Soil = "Arb")
leaf.big <- results(dds, contrast = c("Group", "Leaf.Ino.Big", "Leaf.NC.Big")) %>% tidy() %>% mutate(Tissue = "Leaf", Soil = "Big")
leaf.dav <- results(dds, contrast = c("Group", "Leaf.Ino.Dav", "Leaf.NC.Dav")) %>% tidy() %>% mutate(Tissue = "Leaf", Soil = "Dav")
root.arb <- results(dds, contrast = c("Group", "Root.Ino.Arb", "Root.NC.Arb")) %>% tidy() %>% mutate(Tissue = "Root", Soil = "Arb")
root.big <- results(dds, contrast = c("Group", "Root.Ino.Big", "Root.NC.Big")) %>% tidy() %>% mutate(Tissue = "Root", Soil = "Big")
root.dav <- results(dds, contrast = c("Group", "Root.Ino.Dav", "Root.NC.Dav")) %>% tidy() %>% mutate(Tissue = "Root", Soil = "Dav")

trt.res <- rbind(leaf.arb, leaf.big, leaf.dav,
                 root.arb, root.big, root.dav)

```

Generate DESeq object to test the effect of soil type on plant transcriptomes
```{r}
dds2 <- DESeqDataSetFromTximport(txi,
                                colData = map,
                                design = ~ Soil)
```

Run an LRT testing the effect of soil type
```{r}
root.map <- filter(map, Tissue == "Root" & Treatment == "Ino")
root.dds <-  dds2[, as.vector(root.map$SampleID)]
root.dds <- DESeq(root.dds, test = "LRT", reduced = ~ 1)
root.soil.res <- results(root.dds) %>% tidy() %>% mutate(Tissue = "Root", Treatment = "Ino")

leaf.map <- filter(map, Tissue == "Leaf" & Treatment == "Ino")
leaf.dds <-  dds2[, as.vector(leaf.map$SampleID)]
leaf.dds <- DESeq(leaf.dds, test = "LRT", reduced = ~ 1)
leaf.soil.res <- results(leaf.dds) %>% tidy() %>% mutate(Tissue = "Leaf", Treatment = "Ino")

root.nc.map <- filter(map, Tissue == "Root" & Treatment == "NC")
root.nc.dds <-  dds2[, as.vector(root.nc.map$SampleID)]
root.nc.dds <- DESeq(root.nc.dds, test = "LRT", reduced = ~ 1)
root.nc.soil.res <- results(root.nc.dds) %>% tidy() %>% mutate(Tissue = "Root", Treatment = "NC")

leaf.nc.map <- filter(map, Tissue == "Leaf" & Treatment == "NC")
leaf.nc.dds <-  dds2[, as.vector(leaf.nc.map$SampleID)]
leaf.nc.dds <- DESeq(leaf.nc.dds, test = "LRT", reduced = ~ 1)
leaf.nc.soil.res <- results(leaf.nc.dds) %>% tidy() %>% mutate(Tissue = "Leaf", Treatment = "NC")

soil.res <- rbind(root.soil.res, leaf.soil.res, root.nc.soil.res, leaf.nc.soil.res) 
```

Save results
```{r}
saveRDS(dds, "../Data/msu_dds.RDS")
saveRDS(trt.res, "../Data/msu_trt_res.RDS")
saveRDS(soil.res, "../Data/msu_soil_res.RDS")
```
