Load libraries
```{r}
library(DESeq2)
library(biobroom)
library(ggdendro)
library(vegan)
library(RColorBrewer)
library(cowplot)
library(tidyverse)

source("../../General/rmb_functions.R")
source("../../General/parameters.R")
```

Load data
```{r}
map <- readRDS("../Data/GNOTO_MB_map.RDS") %>% 
  filter(Treatment != "NC") %>% 
  mutate(Set = case_when(Compartment == "BS" ~ "BS",
                         Compartment == "ES" & Treatment == "PC" ~ "ES.PC",
                         Compartment == "ES" & Treatment == "Ino" ~ "ES.Ino"))

otu <- readRDS("../Data/SILVAcr_otu.RDS")
tax <- readRDS("../Data/SILVAcr_tax.RDS")

# Remove low prevalent OTUs (occupancy < 5% samples)
otu <- otu[,colnames(otu) %in% map$SampleID]
otu <- otu[rowSums(otu) > 0.05*ncol(otu),]
```

Get relative abundances and top phyla
```{r}
# Calculate relative abundances
otu.ra <- rel_ab(otu)
otu.ra.tidy <- tidy_otu(otu.ra) %>% mutate(Count = Count/100)

# Get top 11 phyla
tax <- expand_proteo(tax)
top.tax <- get_top_taxa(otu.ra, tax, rank = "PhyClass", n = 11)
tax <- collapse_other(tax, top.tax)

# Pool all other taxa into a "low abundance" category
tax <- tax %>% 
  mutate(PhyClass2 = fct_recode(PhyClass2, "Low abundance" = "other")) %>% 
  mutate(PhyClass2 = fct_relevel(PhyClass2, "Low abundance"))
```

Calculate Bray-Curtis dissimilarities
```{r}
dist <- beta_div_dist(otu.ra %>% log_norm)
```

Perform hierarchical clustering on the Bray-Curtis dissimilarities
```{r}
# Run hierarchical clustring
dd.row <- as.dendrogram(hclust(as.dist(dist), method = "average"))
ddata_x <- dendro_data(dd.row)

# Extract data for plotting dendogram
dendro.p <- ggplot(segment(ddata_x)) +
  geom_segment(aes(x=x, y=y, xend=xend, yend=yend))

# Get the labels for the tips of the dendogram
labs <- label(ddata_x) %>% 
  inner_join(map, by = c("label" = "SampleID")) %>% 
  mutate(Soil = fct_recode(Soil,
                           Arbuckle = "Arb",
                           Biggs = "Big",
                           Davis = "Dav")) %>% 
  mutate(Set = fct_recode(Set,
                          "BS" = "BS.Ino",
                          "ES SI" = "ES.Ino",
                          "ES SG" = "ES.PC")) %>% 
  mutate(Set = fct_relevel(Set, "BS", "ES SG", "ES SI"))

# Generate a data frame with the relative abundances of each phylum
phy.ra <- otu %>%
  tidy_otu() %>% 
  inner_join(tax) %>% 
  group_by(SampleID, PhyClass2) %>%
  summarise(Total = sum(Count)) %>% 
  group_by(SampleID) %>% 
  mutate(RelAb = (Total/sum(Total))) %>% 
  inner_join(map, by = "SampleID") %>% 
  select(SampleID, PhyClass2, RelAb)

phy.ra <- inner_join(labs, phy.ra,  by = c("label" = "SampleID"))

# Plot dendrogram
clad <- dendro.p +
  geom_point(data=labs,
             aes(x=x, y=0.1, shape = Soil, fill = Soil), size = 3, color = "black") +
  scale_fill_brewer(palette = "Greys",
                    guide = guide_legend(title.hjust = 0.5,title.position = "top")) +
  scale_shape_manual(values = c(21,22,24)) +
  theme_classic() +
  xlim(c(0,51)) +
  theme(text = element_text(size = 13),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        legend.position = "top") +
  coord_flip() +
  scale_y_reverse()

# Generate dataframe for plotting group boxes
comps <- phy.ra %>%
  mutate(Group = paste(Compartment, Treatment, sep = ".")) %>%
  mutate(Group = fct_relevel(Group, "BS.Ino", "ES.PC", "ES.Ino")) %>%
  mutate(Group = fct_recode(Group,
                            "Bulk soil" = "BS.Ino",
                            "Endosphere SI" = "ES.Ino",
                            "Endosphere SG" = "ES.PC")) %>%
  group_by(Group) %>%
  summarise(min = min(x),
            max = max(x),
            mid = max(x) - (max(x) - min(x))/2)

# Plot taxonomy barplots
taxplot <- ggplot(phy.ra) +
  geom_bar(aes(x=x, y=RelAb, fill = PhyClass2), stat = "identity") +
  geom_rect(data = comps, aes(xmin = min - 0.4, xmax = max + 0.4, ymin = -0.025, ymax = -0.15, color = Group)) +
  geom_text(data = comps, aes(x = mid, y = -0.085, label = Group, color = Group, angle = 270)) +
  scale_fill_manual(values = phy.pal, name = "Taxon") +
  scale_color_manual(values = c("#6A3D9A", "#FF7F00", "#FDBF6F"), guide = F) +
  theme_classic() +
  xlim(c(0,51)) +
  ylim(c(-0.2,1)) +
  theme(text = element_text(size = 13),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        legend.position = "top") +
  coord_flip() +
  scale_y_reverse()


plot_grid(clad + theme(legend.position = "none"),
                  taxplot + theme(legend.position = "none"),
                  ncol = 2,
                  rel_widths = c(1,2),
                  align = "h",
                  axis = "tb")

```

```{r}
m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")
factoextra::fviz_nbclust(dist, FUN = factoextra::hcut, method = "wss")
# See which method gets you the best clustering
ac <- function(x) {
  cluster::agnes(dist, method = x)$ac
}
map_dbl(m, ac)
```


```{r}
clust.method <- "average"

map.bs <- filter(map, Set == "BS")
map.ei <- filter(map, Set == "ES.Ino")
map.es <- filter(map, Set == "ES.PC")

dist.bs <- dist[match(map.bs$SampleID, row.names(dist)), match(map.bs$SampleID, colnames(dist))]
dist.ei <- dist[match(map.ei$SampleID, row.names(dist)), match(map.ei$SampleID, colnames(dist))]
dist.es <- dist[match(map.es$SampleID, row.names(dist)), match(map.es$SampleID, colnames(dist))]

ddata.bs <- as.dendrogram(hclust(as.dist(dist.bs), method = clust.method)) %>% dendro_data()
labs.bs <- label(ddata.bs) 

ddata.ei <- as.dendrogram(hclust(as.dist(dist.ei), method = clust.method)) %>% dendro_data()
labs.ei <- label(ddata.ei) 

ddata.es <- as.dendrogram(hclust(as.dist(dist.es), method = clust.method)) %>% dendro_data()
labs.es <- label(ddata.es) 

segments.all <- rbind(segment(ddata.bs) %>% mutate(Set = "BS"),
                      segment(ddata.ei) %>% mutate(Set = "ES.Ino"),
                      segment(ddata.es) %>% mutate(Set = "ES.PC")) %>% 
  mutate(Set = fct_recode(Set,
                          "Bulk soil" = "BS",
                          "Endosphere SI" = "ES.Ino",
                          "Endosphere SG" = "ES.PC")) %>%
  mutate(Set = fct_relevel(Set, "Bulk soil", "Endosphere SG", "Endosphere SI"))

labs.all <- rbind(labs.bs,
                  labs.ei,
                  labs.es) %>% 
  dplyr::rename("SampleID" = "label") %>% 
  inner_join(map, by = "SampleID") %>% 
  mutate(Soil = fct_recode(Soil,
                           Arbuckle = "Arb",
                           Biggs = "Big",
                           Davis = "Dav")) %>% 
  mutate(Set = fct_recode(Set,
                          "Bulk soil" = "BS",
                          "Endosphere SI" = "ES.Ino",
                          "Endosphere SG" = "ES.PC")) %>%
  mutate(Set = fct_relevel(Set, "Bulk soil", "Endosphere SG", "Endosphere SI"))


dendro.soil <- segments.all %>% 
  ggplot() +
  geom_segment(aes(x=x, y=y, xend=xend, yend=yend)) +
  geom_point(data=labs.all,
             aes(x=x, y=0.05, shape = Soil, fill = Soil), size = 2, color = "black") +
  scale_fill_brewer(palette = "Set2",
                    guide = guide_legend(title.hjust = 0.5,title.position = "top")) +
  scale_shape_manual(values = c(21,22,24)) +
  theme_bw() +
  theme(text = element_text(size = 11),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank(),
        strip.text = element_text(size = 10, colour = "white"),
        strip.background = element_rect(fill="gray25"),
        legend.position = "left") +
  facet_grid(. ~ Set, scales = "free", space = "free")

dendro.soil
```

```{r}
occupancy.df <- otu.ra.tidy %>% 
  inner_join(map.ei, by = "SampleID") %>% 
  group_by(Soil, OTU_ID) %>% 
  summarise(Occupancy = sum(Count>0),
            MeanAb = mean(Count)) %>% 
  filter(Occupancy > 0)

occupancy.df$OTU_ID %>% unique() %>% length()

core.df <- occupancy.df %>% 
  filter(Occupancy > 6) %>% 
  group_by(OTU_ID) %>% 
  mutate(nSoil = n()) %>% 
  mutate(Core = nSoil == 3)

occupancy.df %>% 
  mutate(Prevalent = Occupancy > 6) %>% 
  group_by(Soil, Prevalent) %>% 
  dplyr::count()

core.ids <- core.df %>% filter(Core) %>% .$OTU_ID %>% unique()
saveRDS(core.ids, "../Data/core_otu_ids.RDS")

occupancy.soil <- occupancy.df %>%
  mutate(Core = OTU_ID %in% filter(core.df, Core)$OTU_ID) %>% 
  ggplot(aes(MeanAb, Occupancy, color = Soil)) +
  geom_point(alpha = 0.2) +
  geom_hline(yintercept = 6.5, linetype = 2) +
  scale_x_log10() +
  scale_color_brewer(palette = "Set2") +
  facet_grid(. ~ Soil) +
  xlab("Mean relative abundance in ES SI") +
  ylab("Occupancy in ES SI\n(# samples)") +
  scale_x_log10(labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  theme_bw() +
  theme(text = element_text(size = 11),
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(size = 10, colour = "white"),
        strip.background = element_rect(fill="gray25"),
        legend.position = "none")


#library(eulerr)

venn.soil <- core.df.venn <- plot(eulerr::venn(list(Arbuckle = filter(core.df, Soil == "Arbuckle")$OTU_ID,
                Biggs = filter(core.df, Soil == "Biggs")$OTU_ID,
                Davis = filter(core.df, Soil == "Davis")$OTU_ID)),
     fills = c("white", "white", "white"),
     edges = RColorBrewer::brewer.pal(3, "Set2"), lex = 3,
     labels = list(fontfamily = "Helvetica",
                   cex = 0),
     quantities = list(fontfamily = "Helvetica",
                   cex = 0.8))


ab.soil <- otu.ra.tidy %>% 
  inner_join(map.ei, by = "SampleID") %>% 
  group_by(Soil, OTU_ID) %>% 
  summarise(AggRelAb = mean(Count)) %>% 
  mutate(Core = OTU_ID %in% filter(core.df, Core)$OTU_ID) %>% 
  filter(OTU_ID %in% filter(core.df, Core)$OTU_ID) %>% 
  group_by(Soil) %>% 
  summarise(TotalRelAb = sum(AggRelAb)) %>% 
  ggplot(aes(Soil, TotalRelAb, fill = Soil)) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette = "Set2") +
  ylab("Relative abundance\nof core OTUs in ES SI") +
  theme_bw() +
  theme(text = element_text(size = 11),
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank(),
        legend.position = "none")


ab.soil



bottom <- plot_grid(occupancy.soil, NA, venn.soil, NA, ab.soil, nrow = 1, rel_widths = c(6,0.5,2,0.5,2.5), align = "h", axis = "b", labels = c("(b)",NA, "(c)",NA,"(d)"), label_size = 15)

#709:408
plot_grid(dendro.soil, bottom, nrow = 2, labels = c("(a)", NA), label_size = 15, rel_heights = c(2,3))
```


```{r}
rank.diff <- otu.ra.tidy %>% 
  group_by(SampleID) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(Compartment == "ES") %>% 
  group_by(Treatment, Soil, OTU_ID) %>% 
  summarise(MeanRelAb = mean(Count)) %>% 
  filter(MeanRelAb > 0) %>% 
  group_by(Treatment, Soil) %>% 
  mutate(Rank = rank(-MeanRelAb, ties.method = "first")) %>% 
  select(-MeanRelAb) %>% 
  spread(key = Treatment, value = Rank) %>% 
  mutate(RankDiff = PC - Ino)
```

```{r}
rank.pc <- otu.ra.tidy %>% 
  group_by(SampleID) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(Compartment == "ES") %>% 
  group_by(Treatment, Soil, OTU_ID) %>% 
  summarise(MeanRelAb = mean(Count)) %>% 
  filter(MeanRelAb > 0) %>% 
  group_by(Treatment, Soil) %>% 
  mutate(Rank = rank(-MeanRelAb, ties.method = "first")) %>% 
  filter(Treatment == "PC") %>% 
  inner_join(select(rank.diff, Soil, OTU_ID, RankDiff), by = c("Soil", "OTU_ID")) %>% 
  mutate(Control = !is.na(RankDiff)) %>% 
  arrange(Soil, Rank) %>% 
  group_by(Soil) %>% 
  mutate(CumAb = cumsum(MeanRelAb))

rank.pc.thr <- rank.pc  %>% 
  filter(CumAb > 0.9) %>% 
  mutate(minRank = min(Rank)) %>% 
  filter(Rank == minRank)

rank.pc.ratio.all <- rank.pc %>% 
  group_by(Soil, Control) %>% 
  dplyr::count() %>% 
  group_by(Soil) %>% 
  mutate(Total = sum(n)) %>% 
  ungroup() %>% 
  filter(Control) %>% 
  mutate(AllRatio = paste(n, Total, sep = " / "))

rank.pc.ratio.top <- rank.pc %>% 
  filter(CumAb <= 0.9) %>% 
  group_by(Soil, Control) %>% 
  dplyr::count() %>% 
  group_by(Soil) %>% 
  mutate(Total = sum(n)) %>% 
  ungroup() %>% 
  filter(Control) %>% 
  mutate(TopRatio = paste(n, Total, sep = " / "))

rank.pc.ratio <- rank.pc.ratio.all %>% 
  select(Soil, AllRatio) %>% 
  inner_join(select(rank.pc.ratio.top, Soil, TopRatio)) %>% 
  mutate(Ratio = paste("All OTUs: ", AllRatio, "\nTop OTUs: ", TopRatio, sep = ""))

rank.pc.p <- rank.pc %>% 
  ggplot() +
  geom_point(aes(Rank, MeanRelAb, color = Control), alpha = 0.5, size = 1) +
  geom_vline(data = rank.pc.thr, aes(xintercept = Rank), linetype = 2) +
  geom_text(data = rank.pc.ratio, aes(x = 500, y = 0.05, label = Ratio), size = 3, hjust = 0) +
  scale_y_log10(labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  scale_color_manual(name = "Detected\nin ES SI", values = c("gray50", "#FDBF6F")) +
  xlab("Abundance rank in ES SG") +
  ylab("Mean relative abundance\nin ES SG") +
  facet_grid(. ~ Soil) +
  theme_bw() +
  theme(text = element_text(size = 11),
        strip.text = element_blank(),
        legend.position = "right") 
  



rank.pc.hist <- rank.pc %>% 
  mutate(Control = !is.na(RankDiff)) %>% 
  ggplot(aes(Rank, fill = Control)) +
  geom_histogram(binwidth = 20, ) +
  geom_vline(data = rank.pc.thr, aes(xintercept = Rank), linetype = 2) +
  scale_fill_manual(name = "Detected\nin ES SI", values = c("gray50", "#FDBF6F")) +
  facet_grid(. ~ Soil) +
  ylab("# of OTUs") +
  theme_bw() +
  theme(text = element_text(size = 11),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_text(size = 10, colour = "white"),
        strip.background = element_rect(fill="gray25"),
        legend.position = "none")


plot_grid(rank.pc.hist, rank.pc.p, nrow =2, align = "v", axis = "lr", rel_heights = c(1,2))
```

```{r}
rank.ino <- otu.ra.tidy %>% 
  group_by(SampleID) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(Compartment == "ES") %>% 
  group_by(Treatment, Soil, OTU_ID) %>% 
  summarise(MeanRelAb = mean(Count)) %>% 
  filter(MeanRelAb > 0) %>% 
  group_by(Treatment, Soil) %>% 
  mutate(Rank = rank(-MeanRelAb, ties.method = "first")) %>% 
  filter(Treatment == "Ino") %>% 
  inner_join(select(rank.diff, Soil, OTU_ID, RankDiff), by = c("Soil", "OTU_ID")) %>% 
  mutate(Control = !is.na(RankDiff)) %>% 
  arrange(Soil, Rank) %>% 
  group_by(Soil) %>% 
  mutate(CumAb = cumsum(MeanRelAb))

rank.ino.thr <- rank.ino  %>% 
  filter(CumAb > 0.9) %>% 
  mutate(minRank = min(Rank)) %>% 
  filter(Rank == minRank)

rank.ino.ratio.all <- rank.ino %>% 
  group_by(Soil, Control) %>% 
  dplyr::count() %>% 
  group_by(Soil) %>% 
  mutate(Total = sum(n)) %>% 
  ungroup() %>% 
  filter(Control) %>% 
  mutate(AllRatio = paste(n, Total, sep = " / "))

rank.ino.ratio.top <- rank.ino %>% 
  filter(CumAb <= 0.9) %>% 
  group_by(Soil, Control) %>% 
  dplyr::count() %>% 
  group_by(Soil) %>% 
  mutate(Total = sum(n)) %>% 
  ungroup() %>% 
  filter(Control) %>% 
  mutate(TopRatio = paste(n, Total, sep = " / "))

rank.ino.ratio <- rank.ino.ratio.all %>% 
  select(Soil, AllRatio) %>% 
  inner_join(select(rank.ino.ratio.top, Soil, TopRatio)) %>% 
  mutate(Ratio = paste("All OTUs: ", AllRatio, "\nTop OTUs: ", TopRatio, sep = ""))

rank.ino.p <- rank.ino %>% 
  ggplot() +
  geom_point(aes(Rank, MeanRelAb, color = Control), alpha = 0.5, size = 1) +
  geom_vline(data = rank.ino.thr, aes(xintercept = Rank), linetype = 2) +
  geom_text(data = rank.ino.ratio, aes(x = 500, y = 0.05, label = Ratio), size = 3, hjust = 0) +
  scale_y_log10(labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  scale_color_manual(name = "Detected\nin ES SG", values = c("gray50", "#FF7F00")) +
  xlab("Abundance rank in ES SI") +
  ylab("Mean relative abundance\nin ES SI") +
  facet_grid(. ~ Soil) +
  theme_bw() +
  theme(text = element_text(size = 11),
        strip.text = element_blank(),
        legend.position = "right")

rank.ino.hist <- rank.ino %>% 
  mutate(Control = !is.na(RankDiff)) %>% 
  ggplot(aes(Rank, fill = Control)) +
  geom_histogram(binwidth = 20, ) +
  geom_vline(data = rank.ino.thr, aes(xintercept = Rank), linetype = 2) +
  scale_fill_manual(name = "Detected\nin ES SG", values = c("gray50", "#FF7F00")) +
  facet_grid(. ~ Soil) +
  ylab("# of OTUs") +
  theme_bw() +
  theme(text = element_text(size = 11),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_text(size = 10, colour = "white"),
        strip.background = element_rect(fill="gray25"),
        legend.position = "none")


plot_grid(rank.ino.hist, rank.ino.p, nrow =2, align = "v", axis = "lr", rel_heights = c(1,2))


```

```{r}
es.map <- map %>% 
  filter(Compartment == "ES") %>% 
  mutate(Group = paste(Treatment, Soil, sep = "."))

es.otu <- otu[,match(es.map$SampleID, colnames(otu))]
es.otu <- es.otu[rowSums(es.otu)>0,]


# Load into DESeq
es.dds <- DESeqDataSetFromMatrix(es.otu,
                                colData = es.map,
                                design = ~ Group)

es.dds <- DESeq(es.dds)

# Run contrasts
arb <- lfcShrink(es.dds, contrast = c("Group", "Ino.Arbuckle", "PC.Arbuckle")) %>% tidy() %>% mutate(Soil = "Arbuckle")
big <- lfcShrink(es.dds, contrast = c("Group", "Ino.Biggs", "PC.Biggs")) %>% tidy() %>% mutate(Soil = "Biggs")
dav <- lfcShrink(es.dds, contrast = c("Group", "Ino.Davis", "PC.Davis")) %>% tidy() %>% mutate(Soil = "Davis")

# Compile and perform adjustments after pooling all contrasts
es.res <- rbind(arb, big, dav) %>% 
  dplyr::rename("OTU_ID" = "gene") %>% 
  ungroup() %>% 
  filter(!is.na(p.adjusted)) %>% 
  mutate(p.adjusted2 = p.adjust(p.value, method = "fdr"))

supp.table <- es.res %>% 
  filter(p.adjusted2 < 0.05) %>% 
  mutate(Direction = ifelse(estimate > 0, "Enriched in SI", "Enriched in SG")) %>% 
  select(Soil, OTU_ID, Direction, everything()) %>% 
  select(-p.adjusted) %>% 
  dplyr::rename("p.adjusted" = "p.adjusted2")

write.table(supp.table, "../Tables/supp.dao.si.vs.sg.tsv", quote = F, sep = "\t", row.names = F, col.names = T)
```


```{r}
cor.df <- otu.ra.tidy %>% 
  group_by(SampleID) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(Compartment == "ES") %>% 
  group_by(Treatment, Soil, OTU_ID) %>% 
  summarise(MeanRelAb = mean(Count)) %>% 
  filter(MeanRelAb > 0) %>% 
  group_by(Treatment, Soil) %>% 
  mutate(Rank = rank(-MeanRelAb, ties.method = "first")) %>% 
  select(-Rank) %>% 
  spread(key = Treatment, value = MeanRelAb) %>% 
  filter(!is.na(Ino) & !is.na(PC)) %>% 
  left_join(es.res, by = c("Soil", "OTU_ID")) %>% 
  mutate(DAO = case_when(p.adjusted2 < 0.05 ~ TRUE,
                         p.adjusted2 >= 0.05 ~ FALSE,
                         is.na(p.adjusted2) ~ FALSE)) 

cor.ratio <- cor.df %>% 
  group_by(Soil, DAO) %>% 
  dplyr::count() %>% 
  group_by(Soil) %>% 
  mutate(Total = sum(n)) %>% 
  filter(DAO) %>% 
  mutate(Ratio = paste(n, Total, sep = " / "))

cor.p <- cor.df %>% 
  ggplot(aes(Ino, PC)) +
  geom_point(alpha = 0.5, aes(color = DAO), size = 1) +
  geom_abline(slope = 1, intercept = 0, linetype = 2) +
  geom_text(data = cor.ratio, aes(x = 0.00001, y = 0.1, label = Ratio), size = 3, hjust = 0) +
  scale_x_log10(labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  scale_y_log10(labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  xlab("Mean relative abundance in ES SI") +
  ylab("Mean relative abundance\nin ES SG") +
  facet_grid(. ~ Soil) +
  scale_color_manual(name = "Differentially\nabundant\nSI vs SG", values = c("gray50", "dodgerblue2")) +
  theme_bw() +
  theme(text = element_text(size = 11),
        strip.text = element_text(size = 10, colour = "white"),
        strip.background = element_rect(fill="gray25"))

cor.p
```
685:916
```{r}
plot_grid(rank.pc.hist, rank.pc.p, rank.ino.hist, rank.ino.p, cor.p, nrow =5, align = "v", axis = "lr", rel_heights = c(1,2,1,2,2), labels = c("(a)", NA, "(b)", NA, "(c)"), label_size = 15)
```



```{r}
otu.ra.tidy %>% 
  group_by(SampleID) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(Compartment == "ES") %>% 
  group_by(Treatment, Soil, OTU_ID) %>% 
  summarise(MeanRelAb = mean(Count)) %>% 
  filter(MeanRelAb > 0) %>% 
  group_by(Treatment, Soil) %>% 
  mutate(Rank = rank(-MeanRelAb, ties.method = "first")) %>% 
  filter(Treatment == "PC") %>% 
  inner_join(select(rank.diff, Soil, OTU_ID, RankDiff), by = c("Soil", "OTU_ID")) %>% 
  mutate(Control = !is.na(RankDiff)) %>% 
  arrange(Soil, Rank) %>% 
  group_by(Soil) %>% 
  mutate(CumAb = cumsum(MeanRelAb)) %>% 
  filter(CumAb > 0.9) %>% 
  mutate(minRank = min(Rank)) %>% 
  filter(Rank == minRank)
```





```{r}
# Run hierarchical clustring
dd.row <- as.dendrogram(hclust(as.dist(dist.bs), method = "average"))
ddata_x <- dendro_data(dd.row)
labs <- label(ddata_x) %>% 
  inner_join(map, by = c("label" = "SampleID"))

# Extract data for plotting dendogram
dendro.p <- ggplot(segment(ddata_x)) +
  geom_segment(aes(x=x, y=y, xend=xend, yend=yend))

# Get the labels for the tips of the dendogram
labs <- label(ddata_x) %>% 
  inner_join(map, by = c("label" = "SampleID")) %>% 
  mutate(Soil = fct_recode(Soil,
                           Arbuckle = "Arb",
                           Biggs = "Big",
                           Davis = "Dav")) %>% 
  mutate(Set = fct_recode(Set,
                          "BS" = "BS.Ino",
                          "ES SI" = "ES.Ino",
                          "ES SG" = "ES.PC")) %>% 
  mutate(Set = fct_relevel(Set, "BS", "ES SG", "ES SI"))

# Generate a data frame with the relative abundances of each phylum
phy.ra <- otu %>%
  tidy_otu() %>% 
  inner_join(tax) %>% 
  group_by(SampleID, PhyClass2) %>%
  summarise(Total = sum(Count)) %>% 
  group_by(SampleID) %>% 
  mutate(RelAb = (Total/sum(Total))) %>% 
  inner_join(map, by = "SampleID") %>% 
  select(SampleID, PhyClass2, RelAb)

phy.ra <- inner_join(labs, phy.ra,  by = c("label" = "SampleID"))

# Plot dendrogram
clad <- dendro.p +
  geom_point(data=labs,
             aes(x=x, y=0.1, shape = Soil, fill = Soil), size = 3, color = "black") +
  scale_fill_brewer(palette = "Greys",
                    guide = guide_legend(title.hjust = 0.5,title.position = "top")) +
  scale_shape_manual(values = c(21,22,24)) +
  theme_classic() +
  xlim(c(0,51)) +
  theme(text = element_text(size = 13),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        legend.position = "top") +
  coord_flip() +
  scale_y_reverse()
```

Run differential abundance analysis testing the impact of soil type on the abundances of individual OTUs
```{r}
# Generate a subset with all the endospheres from soil-inoculated samples
gnot.map <- filter(map, Set == "ES.Ino") 
gnot.otu <- otu[,colnames(otu) %in% gnot.map$SampleID]
gnot.otu <- gnot.otu[,match(gnot.map$SampleID, colnames(gnot.otu))]
gnot.otu <- gnot.otu[rowSums(gnot.otu)>0,]

# Load to DESeq
gnot.dds <- DESeqDataSetFromMatrix(gnot.otu,
                                colData = gnot.map,
                                design = ~ Soil)
gnot.dds <- DESeq(gnot.dds, test="LRT", reduced=~1)

#Format results
gnot.res <- results(gnot.dds) %>% tidy() %>% mutate(Set = "ES.Ino") %>% rename(OTU_ID = "gene")
gnot.sig <- filter(gnot.res, p.adjusted < 0.05)
```

Generate supplementary table
```{r}
supp.table.otu <- gnot.sig %>% inner_join(tax, by = "OTU_ID") %>% 
  select(-Set)

write.table(supp.table.otu, "../Tables/supp.dao.tsv", quote = F, sep = "\t", row.names = F, col.names = T)
```

Generate a data frame with the mean relative abundances (z-score) of each OTU
```{r}
master.means.tidy <- otu %>%
  tidy_otu() %>% 
  inner_join(select(gnot.map, SampleID, Compartment, Treatment, Soil), by = "SampleID") %>% 
  filter(!is.na(Count)) %>% 
  filter(OTU_ID %in% gnot.sig$OTU_ID) %>% 
  group_by(Treatment,OTU_ID) %>% 
  mutate(zscore = (Count - mean(Count))/sd(Count)) %>% 
  group_by(Treatment, Soil, OTU_ID) %>% 
  summarise(MeanZS = mean(zscore)) %>% 
  ungroup()


# Create a matrix to run hierarchical clustering
sig.mtx <- master.means.tidy %>% 
    mutate(Group = paste(Treatment, Soil, sep = ".")) %>% 
    select(Group, MeanZS, OTU_ID) %>% 
    spread(key = Group, value = MeanZS)
sig.mtx <- as.data.frame(sig.mtx) 
rownames(sig.mtx) <- sig.mtx$OTU_ID 
sig.mtx <- sig.mtx[,-1] 
```

Perform hierarchical clustering on the differentiall abundant OTUs
```{r}
n.clust <- 5

sig.dist <- dist(as.matrix(sig.mtx)) 
sig.clust <- hclust(sig.dist, method = "centroid") 
sig.ord.names <- sig.clust$labels[sig.clust$order] 
sig.ord <- data.frame(OTU_ID = sig.ord.names, order = 1:length(sig.ord.names))

sig.cut <- cutree(sig.clust[c(1,2,4)], k = n.clust)
sig.ord$Cluster <- as.factor(sig.cut[sig.ord$OTU_ID])

# Generate a data frame with the order of each OTU based on the hierarchical clustering analysis
sig.ord <- sig.ord %>% 
  mutate(Cluster = paste("C", Cluster, sep = "")) %>% 
  group_by(Cluster) %>% 
  mutate(nOTU = n()) %>% 
  ungroup() %>% 
  mutate(Cluster2 = paste(Cluster, " (", nOTU, " OTUs)", sep = "")) %>% 
  mutate(Cluster = fct_recode(Cluster,
                              "Enriched\nin Biggs" = "C1",
                              "Enriched\nin Arbuckle" = "C2",
                              "Enriched\nin Davis" = "C3",
                              "Depleted\nin Biggs" = "C4",
                              "Depleted\nin Davis" = "C5")) %>% 
  mutate(Cluster = fct_relevel(Cluster, "Enriched\nin Arbuckle"))

full.master <- inner_join(master.means.tidy, sig.ord, by = "OTU_ID") 

# Plot mean abundances in a heatmap
hm <- full.master %>% 
  mutate(Treatment = fct_recode(Treatment,
                                "Endosphere SI" = "Ino",
                                "Endosphere SG" = "PC")) %>% 
  ggplot(aes(Soil, reorder(OTU_ID, order), fill = MeanZS)) +
  geom_tile() +
  scale_fill_gradientn(name = "Abundance\n(z-score)",
                       breaks = c(-1,0,1),
                       colors = (RColorBrewer::brewer.pal(9, "Greys")),
                       guide = guide_colorbar(title.hjust = 1, title.position = "left")) +
  ylab("Differentially abundant OTU") +
  xlab("") +
  facet_grid(Cluster ~ Treatment, scales = "free", space = "free") +
  theme_bw() +
  theme(text = element_text(size = 13),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank(),
        strip.text.y = element_blank(),
        legend.position = "top")
hm
```

Donut plots
```{r}
#Calculate the number of OTUs assigned to each cluster
cluster.total <- sig.ord %>% 
  group_by(Cluster) %>% 
  summarise(nOTU = n()) %>% 
  ungroup() 

# Plot
donut.p <- sig.ord %>% 
  inner_join(tax, by = "OTU_ID") %>% 
  group_by(Cluster, PhyClass2) %>% 
  summarise(Count = n()) %>% 
  group_by(Cluster) %>% 
  mutate(Fraction = Count / sum(Count)) %>% 
  mutate(ymax = cumsum(Fraction),
         nPhy = n()) %>% 
  mutate(ymin = c(0, ymax[1:nPhy - 1])) %>% 
  ggplot() +
  geom_rect(aes(ymax=ymax, ymin=ymin, xmax= 4, xmin= 3, fill= PhyClass2)) +
  geom_text(data = cluster.total, aes(2, 0, label = nOTU), size = 4) +
  scale_fill_manual(name = "Taxon",
                    values = phy.pal) +
  guides(fill = guide_legend(ncol = 4)) +
  coord_polar(theta="y") + 
  xlim(c(2, 4)) +
  facet_wrap(. ~ Cluster, ncol = 1, strip.position = "top") +
  theme_minimal() +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        text = element_text(size = 13),
        strip.text = element_text(size = 11),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank())

donut.p
```

Put plot together
```{r}
fig3.p <- plot_grid(clad + theme(legend.position = "none"),
                  taxplot + theme(legend.position = "none"),
                  hm + theme(legend.position = "none"),
                  donut.p + theme(legend.position = "none"),
                  ncol = 4,
                  rel_widths = c(1,2,1,1),
                   align = "hv",
                   axis = "t",
                  labels = c("A", NA, "B", NA),
                  label_size = 20)

legend.top <- plot_grid(get_legend(clad), get_legend(hm), nrow = 1)
legend.bottom <- plot_grid(get_legend(donut.p), nrow = 1)

fig3.leg <- plot_grid(legend.top,legend.bottom, ncol = 1)

fig3.p ## 676:700
fig3.leg
```

Plot the aggregated relative abundances of each cluster across all soils
```{r}
clust.tax.p <- otu.ra.tidy %>% 
  inner_join(sig.ord, by = "OTU_ID") %>% 
  inner_join(gnot.map, by = "SampleID") %>% 
  group_by(Cluster, Soil, OTU_ID) %>% 
  summarise(MeanRelAb = mean(Count)) %>% 
  inner_join(tax, by = "OTU_ID") %>% 
  group_by(Soil, PhyClass2, Cluster) %>% 
  summarise(MeanRelAb = sum(MeanRelAb)) %>% 
  ggplot(aes(Soil, MeanRelAb, fill = PhyClass2)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = phy.pal, name = "Taxon") +
  facet_grid(. ~ Cluster) +
  #facet_wrap(~ Cluster) +
  ylab("Mean relative abundance") +
  theme_bw() +
  theme(text = element_text(size = 13),
        strip.text = element_text(size = 12, colour = "white"),
        strip.background = element_rect(fill="gray25"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank(),
        legend.position = "right")

clust.tax.p
```
Plot the distribution of abundance ranks across each cluster across all three soils
```{r}
# Calculate the ranks
oc.ab.df <- otu.ra.tidy %>% 
  filter(Count > 0) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(Treatment == "Ino" & Compartment == "ES") %>% 
  group_by(OTU_ID, Soil) %>%
  summarise(MeanAbundance = mean(Count),
            Occupancy = sum(Count > 0)) %>% 
  group_by(Soil) %>% 
  mutate(Rank = rank(-MeanAbundance, ties.method = "first"))

oc.ab.clstr <- oc.ab.df %>% inner_join(sig.ord)

# Plot
rank.p <- oc.ab.df %>% 
  ggplot(aes(Rank)) +
  geom_rug() +
  geom_density(data = filter(oc.ab.clstr, MeanAbundance > 0), aes(color = Cluster), size = 1, alpha = 0.8) +
  geom_rug(data = filter(oc.ab.clstr), aes(color = Cluster)) +
  facet_grid(. ~ Soil, scales = "free") +
  xlab("Mean abundance rank") +
  ylab("Density") +
  scale_color_manual(values = c(brewer.pal(3,"GnBu"), brewer.pal(3, "PuRd")[c(2,3)])) +
  theme_dark() +
  theme(text = element_text(size = 13),
        strip.text = element_text(size = 12, colour = "white"),
        strip.background = element_rect(fill="gray25"))

rank.p
```

Plot supplementary figure
```{r}
plot_grid(rank.p, clust.tax.p,ncol = 1,
          align = "v", axis ="lr", rel_heights = c(4,5),
          labels = c("A", "B"), label_size = 20)
```


